---
title: 'DiffuChannel benchmark: diffusion vs DeepGANnel across all phenotypes'
author: "Barrett-Jolley lab"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document:
    toc: true
    number_sections: false
    md_extensions: -tex_math_dollars
  html_document:
    toc: true
    df_print: paged
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 8, fig.height = 5, dpi = 150)
```

# Purpose

One pass over every phenotype folder. For each one it reads the real seed record,
the DeepGANnel record and the diffusion (DDPM) record, computes the same metric
set on the two synthetic arms, and then pools the five paired comparisons.

Every arm is scored identically.

```{r packages}
required <- c("dplyr", "tidyr", "ggplot2", "readr", "purrr", "arrow","ggh4x",
              "scales", "patchwork", "tibble","Rtsne", "uwot", "kableExtra")
missing  <- required[!vapply(required, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing)) {
  stop("Install first: install.packages(c(\"",
       paste(missing, collapse = "\", \""), "\"))")
}

library(dplyr)
library(tidyr)
library(ggplot2)
library(readr)
library(purrr)
library(arrow)
library(patchwork)
library(uwot)
library(Rtsne)
library(ggh4x)

set.seed(42)
```

# Configuration

```{r config}
CFG <- list(
  # Paths
  pheno_root = "./Phenotypes",   # holds Phenotype_A, Phenotype_B, ...
  batch_dir  = "./batch_out",    # holds Phenotype_A_DDPM.parquet, ...
  out_dir    = "./benchmark_out",

  raw_pattern = "raw",           # substring marking the seed csv (case-insensitive)
  gan_pattern = "gan",           # substring marking the DeepGANnel csv
  only        = NULL,            # e.g. c("Phenotype_A") to run a subset

  # Recording parameters
  sample_hz  = 1e4,              # fallback only; a Time column takes precedence
  filter_hz  = 1e3,

  # Sample rate for generated files:
  #   "seed"  every arm uses the raw record's rate (generated time columns ignored)
  #   "own"   an arm with a time column uses its own rate
  rate_policy  = "seed",

  # Must match training
  seq_len     = 128,
  slow_factor = 64,

  # Analysis settings
  n_windows_mmd = 400,           # O(n^2) in the kernel: keep modest
  n_perm        = 200,
  # Spectral segment length. "auto" is resolved per phenotype (see below); a
  # number is applied to every phenotype and must fit all of them.
  psd_seg       = "auto",
  psd_seg_min   = 256,           # stop below this segment length
  psd_n_seg_min = 20,            # target segments per arm; fewer is flagged
  equalise_noise = TRUE,         # same segment and sample count for every arm
  acf_lag_max   = 200,
  # Conductance level for the noise segments. "auto" picks, per phenotype,
  # closed or open, whichever every arm can support.
  noise_level   = "auto",        # "auto", "closed" or "open"
  ks_subsample  = 2e5,
  # Amplitude-distribution equivalence test
  amp_win       = 1024, # window length (samples). Too short: noisy per-window
                        # histogram. Too long: picks up drift.
  n_windows_amp = 400,  # windows drawn per side
  amp_alpha     = 0.05, # one-sided equivalence significance level (gives a 90% CI)
  equiv_boot    = 500,  # window-pair / dwell bootstrap draws for the floor/arm
                        # KS-distance distributions (amplitude windows and dwell times)
  param_equiv_boot = 300, # block-bootstrap replicates for the two-Gaussian
                        # parametric amplitude equivalence test (EM fit per
                        # replicate, so lower than equiv_boot)
  # Delta for baseline/amplitude: a multiple of the median per-event noise SD
  # (see noise_sd_delta_for_param). Does not shrink with record length.
  noise_sd_k    = 1,    # margin = this many noise SDs
  event_min_len = 30,   # minimum open/closed run length (samples) for a
                        # per-event SD
  trace_secs    = 0.3,

  # Generation block geometry, used only by the block-join (seam) test.
  # Must match the settings diffusion_app.py generated with.
  block_len      = 32768,       # generation block length, samples
  context_len    = 512,         # carried context at block joins, samples
  env_time_secs  = 5,           # seconds of each record drawn in the envelope
                                # time-course plot; envelope metrics use the
                                # whole record
  long_record_secs = 10,        # seconds of each record drawn in the block-join
                                # plot; the join test uses the whole record

  # Optional extras
  do_manifold    = TRUE,        # UMAP / t-SNE, needs uwot + Rtsne, slow
  n_windows_proj = 400,

  proj_win       = 512,

  # ClusterSignificance on the manifold projections (Bioconductor package;
  # chunk skipped if absent). n_perm_cluster permutations per phenotype per
  # projection method.
  do_cluster_test = TRUE,
  n_perm_cluster  = 100,

  cache = FALSE                   # reuse per-phenotype results if present
  #Careful with caching, it can cause real confusion!!
)

dir.create(CFG$out_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(file.path(CFG$out_dir, "cache"), showWarnings = FALSE, recursive = TRUE)

pal <- c("Real" = "#1b1b1b", "Diffusion" = "#c0392b", "GAN" = "#8e44ad",
         "Real floor" = "#7f8c8d", "White noise" = "#bdc3c7")

theme_channel <- function(base_size = 10) {
  theme_classic(base_size = base_size) +
    theme(strip.background = element_blank(),
          strip.text = element_text(face = "bold", hjust = 0),
          legend.position = "top",
          legend.title = element_blank(),
          legend.key.size = unit(0.8, "lines"),
          axis.text = element_text(colour = "black"))
}

# Figures saved as LZW-compressed TIFF. The device is opened directly so the
# compression argument reaches it (ggsave can drop it). ragg if installed
# (better text on Windows), else grDevices::tiff.
save_fig <- function(plot, name, width = 8, height = 5, dpi = 300) {
  path <- file.path(CFG$out_dir, paste0(name, ".tiff"))

  if (requireNamespace("ragg", quietly = TRUE)) {
    ragg::agg_tiff(path, width = width, height = height, units = "in",
                   res = dpi, compression = "lzw", background = "white")
  } else {
    grDevices::tiff(path, width = width, height = height, units = "in",
                    res = dpi, compression = "lzw", bg = "white")
  }
  on.exit(grDevices::dev.off(), add = TRUE)

  print(plot)
  invisible(plot)
}

# kable() scaled to page width in PDF (latex_options ignored in HTML).
# hold_position keeps the table next to its caption.
nice_kable <- function(x, caption = NULL, ...) {
  if (!is.null(caption) && knitr::is_latex_output())
    caption <- gsub("([_$&%#])", "\\\\\\1", caption)
  knitr::kable(x, caption = caption, ...) |>
    kableExtra::kable_styling(latex_options = c("scale_down", "hold_position"))
}
```

# Reading records

The seed and GAN csv files were not written by this pipeline, so their headers
cannot be assumed. Named columns are used when they are recognisable; otherwise
the columns are inferred the same way `batch_run.py` infers them. The resolved
mapping is printed in the manifest below.

```{r fn-read}
normalise_names <- function(nms) gsub("[_-]", " ", trimws(tolower(nms)))

# 10%-of-range quantisation, as in batch_run.py stage_seed()
quantise <- function(x) {
  span <- diff(range(x, na.rm = TRUE))
  if (span > 0) round(x / (0.1 * span)) else rep(0, length(x))
}

# A channels column may hold no more than this many discrete levels (MAX_LEVELS
# in batch_run.py). Used both to infer the column and to check a named one.
MAX_LEVELS <- 10

is_time_col <- function(x) {
  d <- diff(x)
  length(d) > 1 && all(d > 0) &&
    isTRUE(all.equal(d, rep(d[1], length(d)), tolerance = 1e-3))
}

infer_cols <- function(df) {
  # The column count decides whether a time column exists at all: three columns
  # means one of them is time, two means there is none. Which one it is still
  # has to be found, because the order varies.
  time_col <- NA_integer_
  if (ncol(df) >= 3) {
    hit <- which(vapply(df, is_time_col, logical(1)))
    if (!length(hit)) {
      stop(ncol(df), " columns but none of them increases in constant steps, ",
           "so the time column cannot be identified")
    }
    time_col <- hit[1]
  }
  rest <- setdiff(seq_along(df), if (is.na(time_col)) integer(0) else time_col)
  lev  <- lapply(rest, function(j) sort(unique(quantise(df[[j]]))))
  n    <- vapply(lev, length, integer(1))
  k    <- which.min(n)
  if (n[k] > MAX_LEVELS) {
    stop("no column with <= ", MAX_LEVELS, " discrete levels; level counts ",
         paste(n, collapse = ", "))
  }
  others <- setdiff(seq_along(rest), k)
  list(time   = time_col,
       chan   = rest[k],
       cur    = rest[others[which.max(n[others])]],
       levels = lev[[k]])
}

read_record <- function(path) {
  df <- if (grepl("\\.parquet$", path, ignore.case = TRUE)) {
    as.data.frame(arrow::read_parquet(path))
  } else {
    d <- suppressWarnings(as.data.frame(
      readr::read_csv(path, show_col_types = FALSE, name_repair = "minimal")))
    # A file with no header gets its first data row promoted to names; spot that
    # by the names parsing as numbers, and re-read.
    if (all(!is.na(suppressWarnings(as.numeric(names(d)))))) {
      d <- as.data.frame(readr::read_csv(path, col_names = FALSE,
                                         show_col_types = FALSE))
    }
    d
  }

  df <- df[, vapply(df, function(x) is.numeric(x) || is.logical(x) ||
                      !all(is.na(suppressWarnings(as.numeric(as.character(x))))),
                    logical(1)), drop = FALSE]
  df[] <- lapply(df, function(x) suppressWarnings(as.numeric(as.character(x))))
  df <- df[stats::complete.cases(df), , drop = FALSE]
  if (ncol(df) < 2 || nrow(df) < 100) stop("too few usable rows/columns")

  nn  <- normalise_names(names(df))
  idx <- function(target) { h <- which(nn %in% target); if (length(h)) h[1] else NA_integer_ }
  i_chan <- idx(c("channels", "channel", "ideal", "idealisation", "idealization"))
  i_cur  <- idx(c("noisy current", "current", "raw", "raw current", "i"))
  i_time <- idx(c("time", "t"))

  # A recognised channels/current pair is used as named; otherwise columns are
  # inferred. Column order is not assumed.
  if (!is.na(i_chan) && !is.na(i_cur) && i_chan != i_cur) {
    how  <- "named columns"
    chan <- as.integer(round(df[[i_chan]]))
  } else {
    inf <- infer_cols(df)
    i_chan <- inf$chan; i_cur <- inf$cur; i_time <- inf$time
    how  <- "inferred"
    chan <- match(quantise(df[[i_chan]]), inf$levels) - 1L
  }

  out <- tibble::tibble(channels = as.numeric(chan),
                        current  = as.numeric(df[[i_cur]]))
  out$time <- if (!is.na(i_time)) as.numeric(df[[i_time]]) else
    (seq_len(nrow(out)) - 1) / CFG$sample_hz
  out <- dplyr::filter(out, is.finite(channels), is.finite(current))

  attr(out, "map") <- sprintf(
    "%s: chan=col%d cur=col%d time=%s", how, i_chan, i_cur,
    if (is.na(i_time)) sprintf("none, assuming %g Hz", CFG$sample_hz)
    else paste0("col", i_time))
  dplyr::select(out, time, channels, current)
}

fs_of <- function(d, fallback = CFG$sample_hz) {
  if (nrow(d) < 3) return(fallback)
  dt <- stats::median(diff(d$time), na.rm = TRUE)
  if (!is.finite(dt) || dt <= 0) fallback else 1 / dt
}

# First n rows of a record, keeping the "map" attribute (subsetting drops it).
crop_rows <- function(d, n) {
  m <- attr(d, "map")
  d <- d[seq_len(n), ]
  attr(d, "map") <- m
  d
}
```

# Metric functions

```{r fn-metrics}
make_windows <- function(x, win, n_max = Inf, stride = NULL) {
  n <- length(x)
  if (n < win) return(matrix(numeric(0), nrow = 0, ncol = win))
  if (is.null(stride)) stride <- max(1L, floor((n - win) / max(1L, min(n_max, n))))
  starts <- seq(1L, n - win + 1L, by = stride)
  if (length(starts) > n_max) starts <- sample(starts, n_max)
  t(vapply(starts, function(s) x[s:(s + win - 1L)], numeric(win)))
}

sq_dists <- function(A, B) {
  pmax(outer(rowSums(A^2), rowSums(B^2), "+") - 2 * tcrossprod(A, B), 0)
}

# Median heuristic: gamma = 1 / median pairwise squared distance. The fixed 1/L
# used during training underflows to zero on current in pA and pins the
# statistic to its finite-sample floor.
median_gamma <- function(X, Y, n_sub = 200) {
  A <- X[sample(nrow(X), min(n_sub, nrow(X))), , drop = FALSE]
  B <- Y[sample(nrow(Y), min(n_sub, nrow(Y))), , drop = FALSE]
  m <- stats::median(sq_dists(A, B))
  if (!is.finite(m) || m <= 0) 1 / ncol(X) else 1 / m
}

mmd_rbf <- function(X, Y, gamma = NULL, unbiased = TRUE) {
  m <- nrow(X); n <- nrow(Y)
  if (m < 2 || n < 2) return(NA_real_)
  if (is.null(gamma)) gamma <- median_gamma(X, Y)
  Kxx <- exp(-gamma * sq_dists(X, X))
  Kyy <- exp(-gamma * sq_dists(Y, Y))
  Kxy <- exp(-gamma * sq_dists(X, Y))
  if (unbiased) {
    (sum(Kxx) - sum(diag(Kxx))) / (m * (m - 1)) +
      (sum(Kyy) - sum(diag(Kyy))) / (n * (n - 1)) - 2 * mean(Kxy)
  } else {
    mean(Kxx) + mean(Kyy) - 2 * mean(Kxy)
  }
}

mmd_permutation_p <- function(X, Y, n_perm = CFG$n_perm, gamma = NULL) {
  if (is.null(gamma)) gamma <- median_gamma(X, Y)
  obs <- mmd_rbf(X, Y, gamma)
  Z <- rbind(X, Y); m <- nrow(X)
  null <- replicate(n_perm, {
    i <- sample(nrow(Z))
    mmd_rbf(Z[i[seq_len(m)], , drop = FALSE], Z[i[-seq_len(m)], , drop = FALSE], gamma)
  })
  s <- stats::sd(null)
  list(mmd = obs,
       p   = (1 + sum(null >= obs)) / (n_perm + 1),
       z   = if (is.finite(s) && s > 0) (obs - mean(null)) / s else NA_real_)
}

# Noise segments are taken at a single conductance level so no gating step falls
# inside a segment. Runs are kept separate: concatenating them puts a step at
# every join, which adds broadband power and lowers the ACF.
at_level <- function(d, level) {
  if (identical(level, "closed")) d$channels == 0 else d$channels >= max(d$channels)
}

level_runs <- function(d, level, min_len = 1L) {
  r <- rle(at_level(d, level))
  ends <- cumsum(r$lengths); starts <- ends - r$lengths + 1
  keep <- which(r$values & r$lengths >= min_len)
  lapply(keep, function(i) d$current[starts[i]:ends[i]])
}

longest_at <- function(d, level) {
  r <- rle(at_level(d, level))
  if (!any(r$values)) 0L else max(r$lengths[r$values])
}

# Lengths of the runs at one level, without materialising the runs themselves.
# Used to choose a segment length before any spectrum is computed.
run_lengths_at <- function(d, level) {
  r <- rle(at_level(d, level))
  r$lengths[r$values]
}

# Number of Welch segments from a set of run lengths at 50% overlap (as in
# welch_psd).
seg_count <- function(lens, nperseg) {
  lens <- lens[lens >= nperseg]
  if (!length(lens)) return(0L)
  as.integer(sum(1L + (lens - nperseg) %/% (nperseg %/% 2L)))
}

# Welch across a list of runs: every segment from every run long enough to hold
# one contributes to a single average periodogram.
#
# n_seg_max thins the segments to a fixed count, spaced evenly across the record.
# Welch variance falls as 1/n_seg, so both synthetic arms are held to the same
# count.
welch_psd <- function(runs, fs, nperseg, n_seg_max = Inf) {
  if (is.numeric(runs)) runs <- list(runs)
  runs <- runs[vapply(runs, length, integer(1)) >= nperseg]
  if (!length(runs)) stop("no run at the chosen level holds ", nperseg, " samples")

  step <- nperseg %/% 2
  idx <- do.call(rbind, lapply(seq_along(runs), function(i)
    cbind(run = i, start = seq(1L, length(runs[[i]]) - nperseg + 1L, by = step))))
  n_avail <- nrow(idx)

  if (is.finite(n_seg_max) && n_seg_max < n_avail) {
    k <- as.integer(n_seg_max)
    idx <- idx[1L + ((0:(k - 1L)) * n_avail) %/% k, , drop = FALSE]
  }

  w <- 0.5 - 0.5 * cos(2 * pi * (seq_len(nperseg) - 1) / (nperseg - 1))  # Hann
  U <- sum(w^2)
  acc <- numeric(nperseg %/% 2 + 1)
  for (r in seq_len(nrow(idx))) {
    s   <- idx[r, "start"]
    seg <- runs[[idx[r, "run"]]][s:(s + nperseg - 1L)]
    seg <- seg - mean(seg)
    X   <- stats::fft(seg * w)[seq_len(nperseg %/% 2 + 1)]
    acc <- acc + (Mod(X)^2) / (fs * U)
  }
  nseg <- nrow(idx)
  psd <- acc / nseg
  psd[-c(1, length(psd))] <- 2 * psd[-c(1, length(psd))]
  tibble::tibble(freq = seq(0, fs / 2, length.out = length(psd)), psd = psd,
                 n_seg = nseg, n_seg_avail = n_avail) |>
    dplyr::filter(freq > 0)
}

# Length-weighted mean of the per-run ACFs. n_max is a common sample budget per
# arm, spent on the runs in order; the run that crosses it is trimmed.
acf_of <- function(runs, lag_max, fs, n_max = Inf) {
  if (is.numeric(runs)) runs <- list(runs)
  runs <- runs[vapply(runs, length, integer(1)) > lag_max]
  if (!length(runs)) stop("no run at the chosen level is longer than ", lag_max,
                          " samples")

  if (is.finite(n_max)) {
    lens  <- vapply(runs, length, integer(1))
    spent <- c(0L, cumsum(lens))[seq_along(lens)]
    take  <- pmin(lens, pmax(0L, as.integer(n_max) - spent))
    keep  <- take > lag_max
    if (any(keep)) runs <- Map(function(x, k) x[seq_len(k)], runs[keep], take[keep])
  }

  w <- vapply(runs, length, integer(1))
  A <- vapply(runs, function(x)
    stats::acf(x - mean(x), lag.max = lag_max, plot = FALSE,
               demean = TRUE)$acf[, 1, 1], numeric(lag_max + 1))
  tibble::tibble(lag_n  = seq_len(lag_max + 1),
                 lag_ms = 1000 * (seq_len(lag_max + 1) - 1) / fs,
                 acf    = as.numeric(A %*% w / sum(w)),
                 n_used = sum(w))
}

dwell_times <- function(ideal, fs) {
  r <- rle(as.integer(round(ideal)))
  tibble::tibble(level = r$values, n_sample = r$lengths,
                 dwell_ms = 1000 * r$lengths / fs) |>
    dplyr::mutate(state = ifelse(level == 0, "Closed", "Open"))
}

# Half-amplitude idealiser. Two Gaussians of common width, fitted by EM.
# Soft responsibilities, since hard assignment fails when one level holds ~1% of
# the record. Polarity says which of the two fitted levels is closed.
idealise_half_amplitude <- function(current, unit_pA = NULL, baseline = NULL,
                                    n_levels = 1, polarity = 1) {
  x <- current[is.finite(current)]

  if (is.null(baseline) || is.null(unit_pA)) {
    # Fit on a regular subsample of 1e5 points (a 1% level still gives ~1000).
    xf <- if (length(x) > 1e5) x[seq(1, length(x), length.out = 1e5)] else x

    # Start means at the 0.1/99.9 percentiles; starting weight from the split
    # at the midpoint.
    mu  <- stats::quantile(xf, c(0.001, 0.999), names = FALSE)
    sg  <- max(stats::mad(xf), .Machine$double.eps^0.5)
    w   <- min(max(mean(xf < mean(mu)), 1e-3), 1 - 1e-3)
    tol <- 1e-8 * sg
    
    for (i in 1:300) {
      d1 <- stats::dnorm(xf, mu[1], sg) * w
      d2 <- stats::dnorm(xf, mu[2], sg) * (1 - w)
      r  <- d2 / (d1 + d2)
      r[!is.finite(r)] <- 0.5
      s2 <- sum(r); s1 <- sum(1 - r)
      if (s1 < 1 || s2 < 1) break
      new_mu <- c(sum((1 - r) * xf) / s1, sum(r * xf) / s2)
      new_sg <- max(sqrt((sum((1 - r) * (xf - new_mu[1])^2) +
                          sum(r * (xf - new_mu[2])^2)) / length(xf)),
                    .Machine$double.eps^0.5)
      w <- s1 / length(xf)
      done <- max(abs(c(new_mu - mu, new_sg - sg))) < tol
      mu <- new_mu; sg <- new_sg
      if (done) break
    }

    mu   <- sort(mu)
    shut <- if (polarity >= 0) mu[1] else mu[2]
    open <- if (polarity >= 0) mu[2] else mu[1]
    minority <- min(w, 1 - w)

    if (!is.finite(open - shut) || abs(open - shut) < 3 * sg ||
        minority < 5e-4) {
        out <- rep(NA_integer_, length(current))
        attr(out, "why") <- sprintf(
          "sep %.3g, width %.3g, rarer level %.3g",
          abs(open - shut), sg, minority)
        return(out)
    }

    if (is.null(baseline)) baseline <- shut
    if (is.null(unit_pA))  unit_pA  <- open - shut
  }

  lvl <- (current - baseline) / unit_pA
  pmin(pmax(as.integer(round(lvl)), 0L), as.integer(n_levels))
}

idealise_label_anchored_cv <- function(
    current,
    labels,
    n_levels = 1L,
    polarity = 1,
    block_len = 1000L,
    trim = 0.10,
    min_state_n = 50L
) {
  stopifnot(length(current) == length(labels))

  ok <- is.finite(current) & is.finite(labels)
  labels <- as.integer(labels > 0)

  n <- length(current)
  block <- ((seq_len(n) - 1L) %/% block_len) + 1L
  fold <- block %% 2L

  recovered <- rep(NA_integer_, n)
  fit_info <- vector("list", 2L)

  robust_centre <- function(x) {
    x <- x[is.finite(x)]

    if (!length(x)) {
      return(NA_real_)
    }

    mean(x, trim = trim)
  }

  for (test_fold in 0:1) {
    train <- ok & fold != test_fold
    test  <- ok & fold == test_fold

    x_closed <- current[train & labels == 0L]
    x_open   <- current[train & labels > 0L]

    if (length(x_closed) < min_state_n ||
        length(x_open) < min_state_n) {
      next
    }

    closed_centre <- robust_centre(x_closed)
    open_centre   <- robust_centre(x_open)

    if (!is.finite(closed_centre) ||
        !is.finite(open_centre) ||
        closed_centre == open_centre) {
      next
    }

    # Respect the polarity defined from the real record.
    if (polarity >= 0) {
      baseline <- closed_centre
      unit_pA  <- open_centre - closed_centre
    } else {
      baseline <- closed_centre
      unit_pA  <- open_centre - closed_centre
    }

    lvl <- (current[test] - baseline) / unit_pA

    recovered[test] <- pmin(
      pmax(as.integer(round(lvl)), 0L),
      as.integer(n_levels)
    )

    fit_info[[test_fold + 1L]] <- list(
      n_closed = length(x_closed),
      n_open = length(x_open),
      baseline = baseline,
      open_centre = open_centre,
      unit_pA = unit_pA
    )
  }

  if (all(is.na(recovered))) {
    attr(recovered, "why") <-
      "cross-fitted label anchoring lacked enough observations in one or both states"
  }

  attr(recovered, "fit") <- fit_info
  attr(recovered, "method") <- "cross-fitted label-anchored half-amplitude"

  recovered
}
# Cohen's kappa on two logical vectors (corrects for the high chance agreement
# of records that are nearly always open or nearly always closed).
cohen_kappa <- function(a, b) {
  po <- mean(a == b)
  pe <- mean(a) * mean(b) + mean(!a) * mean(!b)
  if (!is.finite(pe) || pe >= 1) NA_real_ else (po - pe) / (1 - pe)
}

block_mean <- function(x, factor) {
  factor <- max(1L, as.integer(factor))
  nb <- length(x) %/% factor
  if (nb < 1) return(mean(x))
  colMeans(matrix(x[seq_len(nb * factor)], nrow = factor))
}

ks_safe <- function(a, b, n_max = CFG$ks_subsample) {
  a <- a[is.finite(a)]; b <- b[is.finite(b)]
  if (length(a) > n_max) a <- sample(a, n_max)
  if (length(b) > n_max) b <- sample(b, n_max)
  suppressWarnings(stats::ks.test(a, b))
}

# Amplitude equivalence: one-sided window-based test on KS distance
# Tests the whole amplitude distribution (spread and shape as well as location).
# Unit of comparison is a window of CFG$amp_win samples. Delta comes from real
# alone: KS D between random pairs of real windows, one from each half of the
# record. KS D is a non-negative distance, so the test is one-sided at
# CFG$amp_alpha (Wellek, 2010): Delta is the (1 - alpha) quantile of the
# real/real floor, and an arm is equivalent if the (1 - alpha) quantile of the
# real/arm distances is below Delta.
amplitude_floor_windows <- function(real_x, win = CFG$amp_win,
                                     n_windows = CFG$n_windows_amp,
                                     B = CFG$equiv_boot,
                                     alpha = CFG$amp_alpha) {
  n <- length(real_x); half <- n %/% 2L
  W1 <- make_windows(real_x[1:half],       win, n_windows)
  W2 <- make_windows(real_x[(half + 1):n], win, n_windows)
  floor_d <- vapply(seq_len(B), function(i) {
    a <- W1[sample.int(nrow(W1), 1), ]
    b <- W2[sample.int(nrow(W2), 1), ]
    ks_safe(a, b)$statistic
  }, numeric(1))
  unname(stats::quantile(floor_d, 1 - alpha))
}

amplitude_equiv_windows <- function(real_x, arm_x, delta,
                                     win = CFG$amp_win,
                                     n_windows = CFG$n_windows_amp,
                                     B = CFG$equiv_boot,
                                     alpha = CFG$amp_alpha) {
  Wr <- make_windows(real_x, win, n_windows)
  Wa <- make_windows(arm_x,  win, n_windows)
  arm_d <- vapply(seq_len(B), function(i) {
    a <- Wr[sample.int(nrow(Wr), 1), ]
    b <- Wa[sample.int(nrow(Wa), 1), ]
    ks_safe(a, b)$statistic
  }, numeric(1))
  upper <- unname(stats::quantile(arm_d, 1 - alpha))
  p <- (1 + sum(arm_d >= delta)) / (length(arm_d) + 1)  # H0: true D >= Delta

  list(upper = upper, p = p, equivalent = upper < delta)
}

# Parametric amplitude equivalence: two-Gaussian EM fit
# Two Gaussians of common width (closed and open levels) fitted to the amplitude
# histogram. Baseline, unitary amplitude, common SD and open weight (fit-based
# Popen) are signed parameters, tested by TOST with a 90% CI (Schuirmann, 1987).
# Same EM as idealise_half_amplitude(), returning the fitted parameters.
fit_two_gaussian_em <- function(x, polarity = 1) {
  x <- x[is.finite(x)]
  n <- length(x)
  xf <- if (n > 1e5) x[seq(1, n, length.out = 1e5)] else x
  nf <- length(xf)

  mu  <- stats::quantile(xf, c(0.001, 0.999), names = FALSE)
  sg  <- max(stats::mad(xf), .Machine$double.eps^0.5)
  w   <- min(max(mean(xf < mean(mu)), 1e-3), 1 - 1e-3)
  tol <- 1e-8 * sg

  for (i in 1:300) {
    d1 <- stats::dnorm(xf, mu[1], sg) * w
    d2 <- stats::dnorm(xf, mu[2], sg) * (1 - w)
    r  <- d2 / (d1 + d2)
    r[!is.finite(r)] <- 0.5
    s2 <- sum(r); s1 <- sum(1 - r)
    if (s1 < 1 || s2 < 1) break
    new_mu <- c(sum((1 - r) * xf) / s1, sum(r * xf) / s2)
    new_sg <- max(sqrt((sum((1 - r) * (xf - new_mu[1])^2) +
                        sum(r * (xf - new_mu[2])^2)) / nf),
                  .Machine$double.eps^0.5)
    w <- s1 / nf
    done <- max(abs(c(new_mu - mu, new_sg - sg))) < tol
    mu <- new_mu; sg <- new_sg
    if (done) break
  }

  mu   <- sort(mu)
  shut <- if (polarity >= 0) mu[1] else mu[2]
  open <- if (polarity >= 0) mu[2] else mu[1]
  minority <- min(w, 1 - w)
  popen_w  <- if (polarity >= 0) 1 - w else w
  ok <- is.finite(open - shut) && abs(open - shut) >= 3 * sg && minority >= 5e-4

  list(baseline = unname(shut), amplitude = unname(open - shut),
       sd = sg, weight = popen_w, ok = ok)
}

# Parametric amplitude equivalence: single-Beta fallback
# For records whose real trace fails two-Gaussian separation (e.g. Phenotype_D,
# ~97% of its bootstrap replicates). One Beta component, which allows skewed
# noise. Reports only "location" and "spread" (fitted mean and SD, pA) under
# their own metric names. Mode is set once per phenotype from the full real
# trace (see "param_mode"), so real and both arms are fitted the same way.
# x is rescaled to (0, 1) with a fixed lo/hi from real's range (padded 1% each
# side), shared by real and both arms; values outside are clipped.
fit_beta_single <- function(x, lo, hi, eps = 1e-4) {
  x <- x[is.finite(x)]
  # Subsample to 1e5 points, as in fit_two_gaussian_em (runs inside a bootstrap
  # loop).
  n <- length(x)
  x <- if (n > 1e5) x[seq(1, n, length.out = 1e5)] else x
  span <- hi - lo
  u <- (x - lo) / span
  u <- pmin(pmax(u, eps), 1 - eps)

  m <- mean(u); v <- max(stats::var(u), 1e-8)
  k  <- m * (1 - m) / v - 1
  a0 <- max(m * k, 1e-2)
  b0 <- max((1 - m) * k, 1e-2)

  nll <- function(par) {
    a <- exp(par[1]); b <- exp(par[2])
    -sum(stats::dbeta(u, a, b, log = TRUE))
  }
  fit <- tryCatch(
    stats::optim(log(c(a0, b0)), nll, method = "Nelder-Mead",
                 control = list(maxit = 500)),
    error = function(e) NULL
  )
  if (is.null(fit) || fit$convergence != 0) {
    return(list(location = NA_real_, spread = NA_real_, ok = FALSE))
  }
  a <- exp(fit$par[1]); b <- exp(fit$par[2])
  mean_u <- a / (a + b)
  sd_u   <- sqrt(a * b / ((a + b)^2 * (a + b + 1)))
  list(location = lo + mean_u * span, spread = sd_u * span, ok = TRUE)
}

# Dispatch to the Gaussian or Beta fit by mode.
fit_amp_component <- function(x, polarity, mode = "gaussian", beta_range = NULL) {
  if (identical(mode, "beta")) return(fit_beta_single(x, beta_range[1], beta_range[2]))
  fit_two_gaussian_em(x, polarity)
}

# Block bootstrap: resample contiguous blocks with replacement to keep local
# autocorrelation. Used for the parametric fit, which needs full-length
# replicates (a single amp_win window rarely holds enough openings).
block_bootstrap_1d <- function(x, block_len, n_out) {
  n <- length(x)
  n_blocks <- ceiling(n_out / block_len)
  starts <- sample.int(n - block_len + 1L, n_blocks, replace = TRUE)
  out <- unlist(lapply(starts, function(s) x[s:(s + block_len - 1L)]))
  out[seq_len(n_out)]
}

dwell_block_len <- function(channels, mult = 10, lo = 50L, hi = 5000L) {
  runs <- rle(channels)$lengths
  as.integer(max(lo, min(hi, mult * mean(runs))))
}

# Three sets of real-only block-bootstrap fits, made once per phenotype and
# shared by both arms and all parameters: f1/f2 give Delta (the floor), fr is
# compared with each arm's bootstrap fits.
param_floor_fits <- function(real_x, real_channels, polarity, mode = "gaussian",
                              beta_range = NULL, B = CFG$param_equiv_boot) {
  bl <- dwell_block_len(real_channels)
  n  <- length(real_x)
  rep_fit <- function()
    fit_amp_component(block_bootstrap_1d(real_x, bl, n), polarity, mode, beta_range)
  list(f1 = replicate(B, rep_fit(), simplify = FALSE),
       f2 = replicate(B, rep_fit(), simplify = FALSE),
       fr = replicate(B, rep_fit(), simplify = FALSE))
}

param_arm_fits <- function(arm_x, real_channels, polarity, mode = "gaussian",
                            beta_range = NULL, B = CFG$param_equiv_boot) {
  # Block length from real's dwell structure, as for Delta.
  bl <- dwell_block_len(real_channels)
  replicate(B, fit_amp_component(block_bootstrap_1d(arm_x, bl, length(arm_x)), polarity,
                                  mode, beta_range),
            simplify = FALSE)
}

# Delta from real alone (f1 vs f2).
param_delta_field <- function(floor_fits, field) {
  ok1 <- vapply(floor_fits$f1, `[[`, logical(1), "ok")
  ok2 <- vapply(floor_fits$f2, `[[`, logical(1), "ok")
  keep <- ok1 & ok2
  n <- sum(keep)
  if (n < 20) return(list(ok = FALSE, n = n))
  val <- function(fits) vapply(fits, `[[`, numeric(1), field)
  floor_diff <- val(floor_fits$f1)[keep] - val(floor_fits$f2)[keep]
  list(ok = TRUE, n = n, delta = unname(stats::quantile(abs(floor_diff), 0.90)))
}

param_equiv_field <- function(floor_fits, arm_fits, field, delta) {
  ok_r <- vapply(floor_fits$fr, `[[`, logical(1), "ok")
  ok_a <- vapply(arm_fits,      `[[`, logical(1), "ok")
  keep <- ok_r & ok_a
  n_ok <- sum(keep)
  if (n_ok < 20) return(list(ok = FALSE, n_ok = n_ok))
  val <- function(fits) vapply(fits, `[[`, numeric(1), field)
  diff_boot <- val(floor_fits$fr)[keep] - val(arm_fits)[keep]
  ci <- unname(stats::quantile(diff_boot, c(0.05, 0.95)))
  p1 <- (1 + sum(diff_boot <= -delta)) / (n_ok + 1)  # H01: true diff <= -Delta
  p2 <- (1 + sum(diff_boot >=  delta)) / (n_ok + 1)  # H02: true diff >= +Delta
  list(ok = TRUE, n_ok = n_ok, ci_lo = ci[1], ci_hi = ci[2], p = max(p1, p2),
       equivalent = ci[1] > -delta && ci[2] < delta)
}

# Dwell-time distribution equivalence
# One-sided KS-distance test on the full open and closed dwell-time
# distributions, bootstrapping the dwell durations directly (Wellek, 2010).
dwell_delta <- function(real_dwell, B = CFG$equiv_boot, alpha = CFG$amp_alpha) {
  n <- length(real_dwell)
  if (n < 20) return(list(ok = FALSE, n = n))
  half <- n %/% 2L
  r1 <- real_dwell[1:half]; r2 <- real_dwell[(half + 1):n]
  floor_d <- vapply(seq_len(B), function(i) {
    a <- sample(r1, length(r1), replace = TRUE)
    b <- sample(r2, length(r2), replace = TRUE)
    ks_safe(a, b)$statistic
  }, numeric(1))
  list(ok = TRUE, n = n, delta = unname(stats::quantile(floor_d, 1 - alpha)))
}

dwell_equiv <- function(real_dwell, arm_dwell, delta,
                         B = CFG$equiv_boot, alpha = CFG$amp_alpha) {
  if (length(arm_dwell) < 20) return(list(ok = FALSE, n = length(arm_dwell)))
  arm_d <- vapply(seq_len(B), function(i) {
    a <- sample(real_dwell, length(real_dwell), replace = TRUE)
    b <- sample(arm_dwell,  length(arm_dwell),  replace = TRUE)
    ks_safe(a, b)$statistic
  }, numeric(1))
  upper <- unname(stats::quantile(arm_d, 1 - alpha))
  p <- (1 + sum(arm_d >= delta)) / (length(arm_d) + 1)
  list(ok = TRUE, upper = upper, p = p, equivalent = upper < delta,
       n_real = length(real_dwell), n_arm = length(arm_dwell))
}

# Noise SD by event
# One SD per open/closed run of at least CFG$event_min_len samples, using the
# same run detection as dwell_times.
noise_sd_by_event <- function(current, channels, state, min_len = CFG$event_min_len) {
  r <- rle(as.integer(round(channels)))
  ends   <- cumsum(r$lengths)
  starts <- ends - r$lengths + 1L
  run_state <- ifelse(r$values == 0, "Closed", "Open")
  keep <- run_state == state & r$lengths >= min_len
  if (!any(keep)) return(numeric(0))
  vapply(which(keep), function(i) stats::sd(current[starts[i]:ends[i]]), numeric(1))
}

# Noise-SD Delta for the parametric baseline/amplitude TOST
# Delta = CFG$noise_sd_k x the real record's pooled per-event noise SD (closed
# events for baseline, open events for amplitude). A bootstrap Delta shrinks as
# the record gets longer; the noise SD converges to a fixed value. Gaussian mode
# only; Beta mode keeps the bootstrap Delta from param_delta_field.
# Pooled SD weights each event by its degrees of freedom (n - 1).
pooled_event_sd <- function(current, channels, state, min_len = CFG$event_min_len) {
  r <- rle(as.integer(round(channels)))
  ends   <- cumsum(r$lengths)
  starts <- ends - r$lengths + 1L
  run_state <- ifelse(r$values == 0, "Closed", "Open")
  keep <- run_state == state & r$lengths >= min_len
  if (!any(keep)) return(list(ok = FALSE, n = 0L))
  idx <- which(keep)
  sds <- vapply(idx, function(i) stats::sd(current[starts[i]:ends[i]]), numeric(1))
  ns  <- r$lengths[idx]
  pooled_var <- sum((ns - 1) * sds^2) / sum(ns - 1)
  list(ok = TRUE, n = length(idx), sd = sqrt(pooled_var))
}

noise_sd_delta_for_param <- function(field, real_x, real_channels,
                                      k = CFG$noise_sd_k) {
  state  <- if (identical(field, "baseline")) "Closed" else "Open"
  pooled <- pooled_event_sd(real_x, real_channels, state)
  if (!pooled$ok || pooled$n < 3) return(list(ok = FALSE, n = pooled$n))
  list(ok = TRUE, n = pooled$n, delta = k * pooled$sd)
}

# Baseline/amplitude from the channels column
# A second estimate alongside the two-Gaussian/Beta fit (which does not use the
# labels): baseline = median(current) where channels == 0, amplitude =
# median(current) where channels > 0, minus baseline, as in the preflight
# inventory. Needs only some closed- and some open-labelled samples.
channels_baseline_amplitude <- function(current, channels) {
  closed <- current[channels == 0]
  open   <- current[channels > 0]
  if (!length(closed) || !length(open)) {
    return(list(ok = FALSE, n_closed = length(closed), n_open = length(open)))
  }
  baseline <- stats::median(closed)
  list(ok = TRUE, n_closed = length(closed), n_open = length(open),
       baseline = baseline, amplitude = stats::median(open) - baseline)
}

# Block bootstrap of current and channels together, keeping each sample paired
# with its state.
paired_block_bootstrap <- function(current, channels, block_len, n_out) {
  n <- length(current)
  n_blocks <- ceiling(n_out / block_len)
  starts <- sample.int(n - block_len + 1L, n_blocks, replace = TRUE)
  idx <- unlist(lapply(starts, function(s) s:(s + block_len - 1L)))[seq_len(n_out)]
  list(current = current[idx], channels = channels[idx])
}

# Real-only block-bootstrap replicates of (current, channels), shared by both
# arms. One set only: Delta comes from noise_sd_delta_for_param.
channels_amp_real_reps <- function(real_x, real_channels, B = CFG$param_equiv_boot) {
  bl <- dwell_block_len(real_channels)
  n  <- length(real_x)
  replicate(B, {
    r <- paired_block_bootstrap(real_x, real_channels, bl, n)
    channels_baseline_amplitude(r$current, r$channels)
  }, simplify = FALSE)
}

channels_amp_arm_reps <- function(arm_x, arm_channels, real_channels,
                                   B = CFG$param_equiv_boot) {
  # Block length from real's dwell structure.
  bl <- dwell_block_len(real_channels)
  n  <- length(arm_x)
  replicate(B, {
    r <- paired_block_bootstrap(arm_x, arm_channels, bl, n)
    channels_baseline_amplitude(r$current, r$channels)
  }, simplify = FALSE)
}

# TOST (90% CI) on the real-vs-arm bootstrap difference of the channels-based
# estimate.
channels_amp_equiv <- function(real_reps, arm_reps, field, delta) {
  ok_r <- vapply(real_reps, `[[`, logical(1), "ok")
  ok_a <- vapply(arm_reps,  `[[`, logical(1), "ok")
  keep <- ok_r & ok_a
  n_ok <- sum(keep)
  if (n_ok < 20) return(list(ok = FALSE, n_ok = n_ok))
  val <- function(reps) vapply(reps, `[[`, numeric(1), field)
  diff_boot <- val(real_reps)[keep] - val(arm_reps)[keep]
  ci <- unname(stats::quantile(diff_boot, c(0.05, 0.95)))
  p1 <- (1 + sum(diff_boot <= -delta)) / (n_ok + 1)
  p2 <- (1 + sum(diff_boot >=  delta)) / (n_ok + 1)
  list(ok = TRUE, n_ok = n_ok, ci_lo = ci[1], ci_hi = ci[2], p = max(p1, p2),
       equivalent = ci[1] > -delta && ci[2] < delta)
}

# Baseline/amplitude, relative margin
# Difference as a fraction of unitary amplitude. The amplitude is recomputed on
# each bootstrap replicate, so its sampling uncertainty enters the CI.
# Delta_rel(field) = k x (real's pooled per-event noise SD for that state) /
# (real's point-estimate amplitude), from real alone.
channels_amp_relative_delta <- function(field, real_x, real_channels,
                                         real_amplitude_pt, k = CFG$noise_sd_k) {
  state  <- if (identical(field, "baseline")) "Closed" else "Open"
  pooled <- pooled_event_sd(real_x, real_channels, state)
  if (!pooled$ok || pooled$n < 3) return(list(ok = FALSE, n = pooled$n))
  if (!is.finite(real_amplitude_pt) || real_amplitude_pt == 0) {
    return(list(ok = FALSE, n = pooled$n))
  }
  list(ok = TRUE, n = pooled$n, delta = k * pooled$sd / abs(real_amplitude_pt))
}

# Uses the replicates from channels_amp_equiv. Relative difference per draw =
# (real field - arm field) / that draw's real amplitude. Draws whose amplitude
# is below guard_frac x the point estimate are discarded and counted.
channels_amp_relative_equiv <- function(real_reps, arm_reps, field, delta,
                                         real_amplitude_pt, guard_frac = 0.1) {
  ok_r  <- vapply(real_reps, `[[`, logical(1), "ok")
  ok_a  <- vapply(arm_reps,  `[[`, logical(1), "ok")
  amp_r <- vapply(real_reps, function(r) if (r$ok) r$amplitude else NA_real_, numeric(1))
  guard <- guard_frac * abs(real_amplitude_pt)
  fit_ok <- ok_r & ok_a
  keep   <- fit_ok & is.finite(amp_r) & abs(amp_r) >= guard
  n_ok <- sum(keep)
  n_discarded_fitfail <- sum(!fit_ok)
  n_discarded_guard    <- sum(fit_ok & (!is.finite(amp_r) | abs(amp_r) < guard))
  if (n_ok < 20) {
    return(list(ok = FALSE, n_ok = n_ok, n_discarded_guard = n_discarded_guard,
                n_discarded_fitfail = n_discarded_fitfail))
  }
  val <- function(reps, f) vapply(reps, `[[`, numeric(1), f)
  rel_diff <- (val(real_reps, field)[keep] - val(arm_reps, field)[keep]) / amp_r[keep]
  ci <- unname(stats::quantile(rel_diff, c(0.05, 0.95)))
  p1 <- (1 + sum(rel_diff <= -delta)) / (n_ok + 1)
  p2 <- (1 + sum(rel_diff >=  delta)) / (n_ok + 1)
  list(ok = TRUE, n_ok = n_ok, n_discarded_guard = n_discarded_guard,
       n_discarded_fitfail = n_discarded_fitfail,
       ci_lo = ci[1], ci_hi = ci[2], p = max(p1, p2),
       equivalent = ci[1] > -delta && ci[2] < delta)
}

# Noise-SD distribution equivalence and Popen equivalence
# One-sided KS-distance tests (Wellek, 2010) on distributions.
# Noise SD: per-event SDs, through dwell_delta/dwell_equiv.
# Popen: one value per window of CFG$amp_win samples, then an ordinary
# bootstrap over the per-window values.
window_popen <- function(channels, win = CFG$amp_win, n_windows = CFG$n_windows_amp) {
  W <- make_windows(channels, win, n_windows)
  if (nrow(W) == 0) return(numeric(0))
  rowMeans(W > 0)
}

popen_delta <- function(real_channels, win = CFG$amp_win, n_windows = CFG$n_windows_amp,
                         B = CFG$equiv_boot, alpha = CFG$amp_alpha) {
  n <- length(real_channels); half <- n %/% 2L
  # Windows from each half of real, so the two vectors share no window.
  p1 <- window_popen(real_channels[1:half],       win, n_windows)
  p2 <- window_popen(real_channels[(half + 1):n], win, n_windows)
  if (length(p1) < 20 || length(p2) < 20) {
    return(list(ok = FALSE, n = min(length(p1), length(p2))))
  }
  floor_d <- vapply(seq_len(B), function(i) {
    a <- sample(p1, length(p1), replace = TRUE)
    b <- sample(p2, length(p2), replace = TRUE)
    ks_safe(a, b)$statistic
  }, numeric(1))
  list(ok = TRUE, n = length(p1) + length(p2),
       delta = unname(stats::quantile(floor_d, 1 - alpha)))
}

popen_equiv <- function(real_channels, arm_channels, delta,
                         win = CFG$amp_win, n_windows = CFG$n_windows_amp,
                         B = CFG$equiv_boot, alpha = CFG$amp_alpha) {
  pr <- window_popen(real_channels, win, n_windows)
  pa <- window_popen(arm_channels,  win, n_windows)
  if (length(pr) < 20 || length(pa) < 20) return(list(ok = FALSE, n = length(pa)))
  arm_d <- vapply(seq_len(B), function(i) {
    a <- sample(pr, length(pr), replace = TRUE)
    b <- sample(pa, length(pa), replace = TRUE)
    ks_safe(a, b)$statistic
  }, numeric(1))
  upper <- unname(stats::quantile(arm_d, 1 - alpha))
  p <- (1 + sum(arm_d >= delta)) / (length(arm_d) + 1)
  list(ok = TRUE, upper = upper, p = p, equivalent = upper < delta,
       n_real = length(pr), n_arm = length(pa))
}

# Window for the example traces, as _best_window in batch_run.py: a stretch
# with Popen in [lo, hi], else the one nearest the middle of that band.
best_window <- function(chan, n, lo = 0.4, hi = 0.8) {
  if (length(chan) <= n) return(1L)
  cs <- c(0, cumsum(as.numeric(chan)))
  k  <- (cs[(n + 1):length(cs)] - cs[1:(length(cs) - n)]) / n
  inside <- which(k >= lo & k <= hi)
  if (length(inside)) return(inside[ceiling(length(inside) / 2)])
  which.min(abs(k - 0.5 * (lo + hi)))
}

rms_diff <- function(a, b) sqrt(mean((a - b)^2))

# Two-component exponential mixture fitted to dwell times by EM. Falls back to
# one component if the slow/fast ratio is under min_ratio or a component holds
# almost no area (as in fit_tcrit). Dwells only a few samples long (e.g. 0.39 ms
# at 10 kHz) also give one component.
fit_exp_mixture <- function(x, n_iter = 500, tol = 1e-10, min_n = 20L,
                            min_ratio = 3, min_area = 0.02) {
  x <- x[is.finite(x) & x > 0]

  one <- function() tibble::tibble(
    n_components = 1L,
    tau_fast_ms  = if (length(x)) mean(x) else NA_real_,
    tau_slow_ms  = NA_real_,
    area_fast    = NA_real_,
    mean_ms      = if (length(x)) mean(x) else NA_real_,
    n_events     = length(x))

  if (length(x) < min_n) return(one())

  tau <- as.numeric(stats::quantile(x, c(0.25, 0.9), names = FALSE))
  if (!(tau[2] > tau[1])) return(one())
  a <- c(0.5, 0.5)
  ll_old <- -Inf

  for (i in seq_len(n_iter)) {
    d <- cbind(a[1] * stats::dexp(x, 1 / tau[1]),
               a[2] * stats::dexp(x, 1 / tau[2]))
    s <- rowSums(d)
    s[s <= 0] <- .Machine$double.xmin
    w  <- d[, 1] / s
    sw <- sum(w)
    if (!is.finite(sw) || sw < 1 || (length(x) - sw) < 1) return(one())
    a   <- c(sw / length(x), 1 - sw / length(x))
    tau <- c(sum(w * x) / sw, sum((1 - w) * x) / (length(x) - sw))
    if (any(!is.finite(tau)) || any(tau <= 0)) return(one())
    ll <- sum(log(s))
    if (is.finite(ll) && is.finite(ll_old) &&
        abs(ll - ll_old) < tol * abs(ll)) break
    ll_old <- ll
  }

  o <- order(tau); tau <- tau[o]; a <- a[o]
  if (tau[2] / tau[1] < min_ratio || min(a) < min_area) return(one())

  tibble::tibble(n_components = 2L, tau_fast_ms = tau[1],
                 tau_slow_ms = tau[2], area_fast = a[1],
                 mean_ms = mean(x), n_events = length(x))
}
```

# Find the files

```{r discover}
find_one <- function(folder, pattern) {
  f <- list.files(
    folder,
    pattern = "\\.(csv|parquet)$",
    full.names = TRUE,
    ignore.case = TRUE
  )

  hit <- f[
    grepl(pattern, basename(f), ignore.case = TRUE)
  ]

  if (!length(hit)) return(NA_character_)

  if (length(hit) > 1) {
    warning(
      folder, ": ", length(hit),
      " files match '", pattern,
      "', using ", basename(hit[1])
    )
  }

  hit[1]
}

folders <- list.dirs(CFG$pheno_root, recursive = FALSE)
if (!is.null(CFG$only)) folders <- folders[basename(folders) %in% CFG$only]
if (!length(folders)) stop("No phenotype folders in ", CFG$pheno_root)

parquets <- list.files(CFG$batch_dir, pattern = "\\.parquet$", full.names = TRUE)

manifest <- purrr::map_dfr(folders, function(fd) {
  tag <- gsub(" ", "_", basename(fd))
  ddpm <- parquets[grepl(paste0("^", tag), basename(parquets))]
  if (!length(ddpm)) ddpm <- parquets[grepl(tag, basename(parquets), fixed = TRUE)]
  tibble::tibble(
    phenotype = tag,
    raw  = find_one(fd, CFG$raw_pattern),
    gan  = find_one(fd, CFG$gan_pattern),
    ddpm = if (length(ddpm)) ddpm[1] else NA_character_
  )
})

nice_kable(
  manifest |> dplyr::mutate(dplyr::across(c(raw, gan, ddpm), basename)),
  caption = "Manifest: seed, GAN and DDPM file for each phenotype."
)

bad <- manifest |> dplyr::filter(is.na(raw) | is.na(ddpm))
if (nrow(bad)) {
  warning("Dropping phenotypes with no seed or no DDPM file: ",
          paste(bad$phenotype, collapse = ", "))
  manifest <- dplyr::filter(manifest, !is.na(raw), !is.na(ddpm))
}
if (!nrow(manifest)) stop("Nothing left to analyse.")
```

# Pre-flight

Every record is read and checked before any metric is computed; a record that
cannot support the analysis stops the run here.

All records are held in memory for the rest of the notebook. `CFG$only` runs a
subset of phenotypes.

```{r preflight}
recs <- purrr::map(seq_len(nrow(manifest)), function(i) {
  row <- manifest[i, ]
  out <- list(Real = read_record(row$raw), Diffusion = read_record(row$ddpm))
  if (!is.na(row$gan)) out$GAN <- read_record(row$gan)
  out
})
names(recs) <- manifest$phenotype

n_before <- purrr::imap_dfr(recs, function(arms, ph)
  tibble::tibble(phenotype = ph, arm = names(arms),
                 n_read = vapply(arms, nrow, integer(1))))

# Seed record inventory, at full length (before cropping).
seed_inventory <- purrr::imap_dfr(recs, function(arms, ph) {
  d  <- arms$Real
  fs <- fs_of(d)
  tibble::tibble(
    phenotype  = ph,
    seed_file  = basename(manifest$raw[manifest$phenotype == ph]),
    n_rows     = nrow(d),
    sample_khz = round(fs / 1000, 3),
    duration_s = nrow(d) / fs,
    max_levels = max(d$channels),
    popen      = mean(d$channels > 0),
    mean_pA    = mean(d$current),
    sd_pA      = stats::sd(d$current))
})

nice_kable(seed_inventory, digits = 4, label = "seed_inventory",
           caption = paste("Seed records as read, before any",
                           "cropping. n_rows is the number of usable rows",
                           "read from each file."))

readr::write_csv(seed_inventory, file.path(CFG$out_dir, "table1_seed_data.csv"))
```

```{r preflight-crop}
# Full-length diffusion records for the block-join test (cropping would drop
# the later joins). Used only by the seam figures.
recs_full_diffusion <- purrr::map(recs, function(arms) arms$Diffusion)

# GAN and Diffusion are cropped to the shorter of the two (in rows), since
# several metrics depend on length (MMD z, permutation p, KS, spectral
# distances) and the two arms are compared head to head. Real keeps its full
# length: it is the fixed reference, and the equivalence margins come from it.
# In practice the GAN length sets arm_min_len.
for (ph in names(recs)) {
  arms <- recs[[ph]]
  sim_names <- setdiff(names(arms), "Real")
  arm_min_len <- Inf
  for (a in sim_names) { arm_min_len <- min(arm_min_len, nrow(arms[[a]])) }
  for (a in sim_names) { arms[[a]] <- crop_rows(arms[[a]], arm_min_len) }
  recs[[ph]] <- arms
}

crop_report <- n_before |>
  dplyr::left_join(
    purrr::imap_dfr(recs, function(arms, ph)
      tibble::tibble(phenotype = ph, arm = names(arms),
                     n_used = vapply(arms, nrow, integer(1)))),
    by = c("phenotype", "arm")) |>
  dplyr::mutate(kept = n_used / n_read)

nice_kable(crop_report, digits = 3, label = "crop_report",
             caption = paste("Rows read and rows retained. Real keeps its full length;",
                              "GAN and Diffusion are cropped to the shorter of the two."))
readr::write_csv(crop_report, file.path(CFG$out_dir, "crop_report.csv"))
```

```{r preflight-inventory}
inventory <- purrr::imap_dfr(recs, function(arms, ph) {
  purrr::imap_dfr(arms, function(d, a) {
    closed <- d$current[d$channels == 0]
    open   <- d$current[d$channels > 0]
    tibble::tibble(
      phenotype = ph, arm = a, n = nrow(d),
      has_time = !grepl("time=none", attr(d, "map")),
      fs = fs_of(d),
      longest_closed = longest_at(d, "closed"),
      longest_open   = longest_at(d, "open"),
      popen = mean(d$channels > 0), max_level = max(d$channels),
      baseline = median(closed),
      unit_amp = median(open) - median(closed),
      cols = attr(d, "map"))
  })
})

# Catch arms that were min-max scaled or are in the wrong units (unitary
# amplitude more than 3x from the seed's).
scale_check <- inventory |>
  dplyr::group_by(phenotype) |>
  dplyr::mutate(rel = unit_amp / unit_amp[arm == "Real"]) |>
  dplyr::ungroup() |>
  dplyr::filter(arm != "Real", rel < 1 / 3 | rel > 3)
if (nrow(scale_check)) {
  stop("Unitary amplitude differs from the seed by more than 3x in: ",
       paste(sprintf("%s/%s (%.3g vs %.3g)", scale_check$phenotype,
                     scale_check$arm, scale_check$unit_amp,
                     scale_check$unit_amp / scale_check$rel), collapse = "; "),
       ". Likely scaling or units mismatch.")
}

nice_kable(inventory, digits = 4, label = "preflight_inventory",
             caption = "Pre-flight inventory of every record analysed.")
readr::write_csv(inventory, file.path(CFG$out_dir, "preflight_inventory.csv"))
```

## Sample rate

Only the seed record's time column reflects a real acquisition.
`diffusion_app.py` writes its export time as `arange(n) * d * SAMPLE_DT` with
`SAMPLE_DT` fixed at 1e-4, so a DDPM record always claims 10 kHz whatever it was
trained on. A GAN record's time column is whatever DeepGANnel wrote. Where it
disagrees with the seed:

- `rate_policy = "seed"`: the time column is ignored and the record takes the
  seed's rate.
- `rate_policy = "own"`: the record's own rate is used (half the bandwidth,
  events twice as long in time).

The lag-1 autocorrelation of the single-level noise, reported per arm below, is
computed per sample and does not depend on the rate label. An arm close to its
seed was produced at the seed's rate; one well below it was produced at a lower
rate.

```{r rate}
seed_fs <- inventory |>
  dplyr::filter(arm == "Real") |>
  dplyr::mutate(fs = dplyr::if_else(has_time, fs, CFG$sample_hz))
seed_fs <- setNames(seed_fs$fs, seed_fs$phenotype)

rates_under <- function(policy) {
  inventory |>
    dplyr::transmute(
      phenotype, arm,
      fs = if (identical(policy, "own")) {
        dplyr::if_else(has_time, fs, unname(seed_fs[phenotype]))
      } else {
        unname(seed_fs[phenotype])
      })
}

rate_map <- function(policy) {
  r <- rates_under(policy)
  split(setNames(r$fs, r$arm), r$phenotype)
}

disputed <- rates_under("own") |>
  dplyr::rename(fs_own = fs) |>
  dplyr::left_join(dplyr::rename(rates_under("seed"), fs_seed = fs),
                   by = c("phenotype", "arm")) |>
  dplyr::filter(abs(fs_own - fs_seed) / fs_seed > 0.01)

nice_kable(rates_under("seed") |>
               tidyr::pivot_wider(names_from = arm, values_from = fs),
             digits = 1, label = "rate_policy_seed",
             caption = "Sample rate per arm under the seed policy (Hz).")
```

```{r rate-disputed}
if (nrow(disputed)) {
  nice_kable(disputed, digits = 1, label = "rate_policy_disputed",
               caption = "Arms whose sample rate differs between the seed and own policies.")
} else {
  cat("No arm's own time column disagrees with its seed; the two policies coincide.\n")
}

FS_MAIN <- rate_map(CFG$rate_policy)

# Opening direction per phenotype, from the real record's labels, applied to
# every arm.
POL_MAIN <- inventory |>
  dplyr::filter(arm == "Real") |>
  dplyr::transmute(phenotype, polarity = sign(unit_amp))
POL_MAIN <- setNames(POL_MAIN$polarity, POL_MAIN$phenotype)

if (!all(POL_MAIN %in% c(-1, 1))) {
  stop("No measurable unitary amplitude in real record: ",
       paste(names(POL_MAIN)[!(POL_MAIN %in% c(-1, 1))], collapse = ", "))
}
```


Decimation decorrelates adjacent samples, so a record at half the rate has a
lower lag-1 autocorrelation than its seed.

## Noise level and segment length for the spectral metrics

The noise metrics need contiguous samples at one conductance level. A quiet
channel has long closed periods; a record at Popen 0.99 has long open periods
and closures only a sample or two wide. The level is chosen per phenotype as
whichever every arm of that phenotype can support.

An open-level spectrum also has no gating transition inside a segment. It
includes open-channel noise on top of instrumentation noise, the same in all
arms.

Segment length is also set per phenotype; the three arms of a phenotype share
it. `"auto"` takes the largest power of two that gives every arm
`psd_n_seg_min` segments (Welch variance falls as 1/n_seg), and falls back
towards the floor when no length does.

```{r level-and-seg}
cap_at <- inventory |>
  dplyr::group_by(phenotype) |>
  dplyr::summarise(closed = min(longest_closed), open = min(longest_open),
                   .groups = "drop")

levels_used <- cap_at |>
  dplyr::mutate(
    level = if (identical(CFG$noise_level, "auto")) {
      dplyr::if_else(open > closed, "open", "closed")
    } else {
      CFG$noise_level
    },
    cap = dplyr::if_else(level == "open", open, closed))

if (any(levels_used$cap == 0)) {
  z <- levels_used |> dplyr::filter(cap == 0)
  stop("No usable ", paste(unique(z$level), collapse = "/"), " segments at all in: ",
       paste(z$phenotype, collapse = ", "),
       ".")
}

nice_kable(levels_used, digits = 0, label = "levels_used",
             caption = "Longest run every arm can supply at each level, the level used, and the segment-length ceiling (cap).")
```

```{r seg-length-table}
LEVEL_OF <- setNames(levels_used$level, levels_used$phenotype)
CAP_OF   <- setNames(levels_used$cap,   levels_used$phenotype)

# Run lengths at the chosen level, per arm, per phenotype.
lens_of <- purrr::imap(recs, function(arms, ph)
  lapply(arms, run_lengths_at, level = LEVEL_OF[[ph]]))

seg_counts_at <- function(lens, nperseg)
  vapply(lens, seg_count, integer(1), nperseg = nperseg)

# Minimum segment length: psd_seg_min, and longer than the ACF's maximum lag.
SEG_FLOOR <- as.integer(2^ceiling(log2(max(CFG$psd_seg_min, CFG$acf_lag_max + 1))))

pick_nperseg <- function(lens, cap) {
  hi <- 2^floor(log2(cap))
  if (hi < SEG_FLOOR) return(NA_integer_)
  for (nps in 2^(floor(log2(hi)):floor(log2(SEG_FLOOR)))) {
    if (min(seg_counts_at(lens, nps)) >= CFG$psd_n_seg_min) return(as.integer(nps))
  }
  SEG_FLOOR
}

if (identical(CFG$psd_seg, "auto")) {
  NPERSEG_OF <- vapply(names(recs),
                       function(ph) pick_nperseg(lens_of[[ph]], CAP_OF[[ph]]),
                       integer(1))
  starved <- names(NPERSEG_OF)[is.na(NPERSEG_OF)]
  if (length(starved)) {
    stop(paste(starved, collapse = ", "), ": no arm supports ", SEG_FLOOR,
         " contiguous samples at the chosen level (caps ",
         paste(CAP_OF[starved], collapse = ", "), ").")
  }
} else {
  nps <- as.integer(CFG$psd_seg)
  if (nps < CFG$psd_seg_min) {
    stop("CFG$psd_seg = ", nps, " is below CFG$psd_seg_min = ", CFG$psd_seg_min, ".")
  }
  if (nps <= CFG$acf_lag_max) {
    stop("CFG$psd_seg = ", nps, " is not longer than CFG$acf_lag_max = ",
         CFG$acf_lag_max, ".")
  }
  short <- names(CAP_OF)[CAP_OF < nps]
  if (length(short)) {
    stop("CFG$psd_seg = ", nps, " but ", paste(short, collapse = ", "),
         " supports only ", paste(CAP_OF[short], collapse = ", "),
         " samples at the chosen level.")
  }
  NPERSEG_OF <- setNames(rep(nps, length(recs)), names(recs))
}

seg_tbl <- purrr::imap_dfr(lens_of, function(lens, ph) {
  n <- seg_counts_at(lens, NPERSEG_OF[[ph]])
  tibble::tibble(
    phenotype = ph, level = LEVEL_OF[[ph]], cap = CAP_OF[[ph]],
    nperseg = NPERSEG_OF[[ph]],
    seg_ms  = 1000 * NPERSEG_OF[[ph]] / FS_MAIN[[ph]][["Real"]],
    res_hz  = FS_MAIN[[ph]][["Real"]] / NPERSEG_OF[[ph]],
    Real = unname(n["Real"]), Diffusion = unname(n["Diffusion"]),
    GAN = if ("GAN" %in% names(n)) unname(n["GAN"]) else NA_integer_,
    n_used = min(n))
})

nice_kable(seg_tbl, digits = 1, label = "seg_length",
             caption = paste("Segment length chosen per phenotype, and the",
                             "segments each arm can supply at it. n_used is",
                             "the minimum across arms; with equalise_noise,",
                             "GAN and Diffusion are held to their shared",
                             "minimum and Real uses all its segments."))
readr::write_csv(seg_tbl, file.path(CFG$out_dir, "psd_segments.csv"))

thin <- seg_tbl |> dplyr::filter(n_used < CFG$psd_n_seg_min)
if (nrow(thin)) {
  warning("Fewer than CFG$psd_n_seg_min = ", CFG$psd_n_seg_min,
          " segments per arm in: ",
          paste(sprintf("%s (%d)", thin$phenotype, thin$n_used), collapse = ", "),
          ". Spectra are noisy; flagged as psd_n_seg_ok in the provenance table.")
}

if (dplyr::n_distinct(levels_used$level) > 1 ||
    dplyr::n_distinct(NPERSEG_OF) > 1) {
  cat("Level and/or segment length differ between phenotypes; noise metrics\n",
      "are not comparable in absolute terms across phenotypes.\n", sep = "")
}
```
# Per-phenotype analysis

Each phenotype is reduced to a long table of metrics, plus the curves for the
pooled figures. `direction` records which way is better for each metric.

```{r analyse}
analyse_one <- function(row, arms_all, fs, level, nperseg) {
  ph <- row$phenotype

  real    <- arms_all$Real
  arms    <- arms_all[setdiff(names(arms_all), "Real")]
  has_gan <- "GAN" %in% names(arms)

  # Windows and MMD
  W <- list(Real = make_windows(real$current, CFG$seq_len, CFG$n_windows_mmd))
  for (a in names(arms)) W[[a]] <- make_windows(arms[[a]]$current, CFG$seq_len,
                                                CFG$n_windows_mmd)
  half <- floor(nrow(real) / 2)
  W_r1 <- make_windows(real$current[1:half], CFG$seq_len, CFG$n_windows_mmd)
  W_r2 <- make_windows(real$current[(half + 1):nrow(real)], CFG$seq_len, CFG$n_windows_mmd)
  W_wn <- make_windows(rnorm(nrow(real), mean(real$current), sd(real$current)),
                       CFG$seq_len, CFG$n_windows_mmd)

  # One bandwidth for every comparison within the phenotype.
  gam <- median_gamma(W$Real, W$Diffusion)

  # Single-level noise
  # Runs kept separate; every record was checked earlier to hold at least one
  # segment.
  runs <- lapply(arms_all, level_runs, level = level, min_len = nperseg)

  # An arm that gates more often has shorter runs and fewer segments, so
  # simulated arms are held to the segment/sample counts of the thinnest.
  lens       <- lapply(runs, function(x) vapply(x, length, integer(1)))
  seg_avail  <- vapply(lens, seg_count, integer(1), nperseg = nperseg)
  samp_avail <- vapply(lens, function(L) sum(L[L > CFG$acf_lag_max]), integer(1))

  # GAN and Diffusion share a common cap; Real is the reference and is not
  # capped.
  sim_arm_names <- setdiff(names(arms_all), "Real")
  seg_cap_sim  <- if (isTRUE(CFG$equalise_noise)) min(seg_avail[sim_arm_names])  else Inf
  samp_cap_sim <- if (isTRUE(CFG$equalise_noise)) min(samp_avail[sim_arm_names]) else Inf
  seg_cap  <- setNames(ifelse(names(seg_avail)  == "Real", Inf, seg_cap_sim),  names(seg_avail))
  samp_cap <- setNames(ifelse(names(samp_avail) == "Real", Inf, samp_cap_sim), names(samp_avail))

  psd <- purrr::imap_dfr(runs, function(x, a)
    dplyr::mutate(welch_psd(x, fs[[a]], nperseg, seg_cap[[a]]), source = a))
  acf_tbl <- purrr::imap_dfr(runs, function(x, a)
    dplyr::mutate(acf_of(x, CFG$acf_lag_max, fs[[a]], samp_cap[[a]]), source = a))

  # Lag-1 autocorrelation of the single-level noise, per arm. A record at a
  # lower rate than assumed sits well below its seed.
  lag1 <- acf_tbl |> dplyr::filter(lag_n == 2)
  lag1 <- setNames(lag1$acf, lag1$source)

  # Arms at different rates have different frequency and lag grids, so both
  # comparisons interpolate onto a shared axis (a no-op when the grids match).
  curve <- function(tbl, a, xvar, yvar) {
    d <- tbl[tbl$source == a, ]
    list(x = d[[xvar]], y = d[[yvar]])
  }

  psd_rms <- function(a) {
    r <- curve(psd, "Real", "freq", "psd")
    g <- curve(psd, a, "freq", "psd")
    hi <- min(CFG$filter_hz, max(r$x), max(g$x))
    lo <- max(min(r$x), min(g$x))
    grid <- exp(seq(log(lo), log(hi), length.out = 512))
    # rule = 2: the log/exp round trip can put the grid ends ~1 ulp outside the
    # data range, where approx() would return NA.
    rms_diff(approx(r$x, log10(r$y), grid, rule = 2)$y,
             approx(g$x, log10(g$y), grid, rule = 2)$y)
  }

  acf_rms <- function(a) {
    r <- curve(acf_tbl, "Real", "lag_ms", "acf")
    g <- curve(acf_tbl, a, "lag_ms", "acf")
    grid <- seq(0, min(max(r$x), max(g$x)), length.out = CFG$acf_lag_max + 1)
    rms_diff(approx(r$x, r$y, grid, rule = 2)$y, approx(g$x, g$y, grid, rule = 2)$y)
  }

  # Slow envelope
  env <- list(Real = block_mean(real$current, CFG$slow_factor))
  for (a in names(arms)) env[[a]] <- block_mean(arms[[a]]$current, CFG$slow_factor)
  We <- lapply(env, function(e) make_windows(e, min(64, length(e)), CFG$n_windows_mmd))

  # Dwell times
  dw <- purrr::imap_dfr(arms_all, function(d, a)
    dplyr::mutate(dwell_times(d$channels, fs[[a]]), source = a))
  mean_dwell <- function(a, s) {
    v <- dw |> dplyr::filter(source == a, state == s) |> dplyr::pull(dwell_ms)
    if (!length(v)) stop(ph, "/", a, ": no ", s, " dwells to average")
    mean(v)
  }

  level_pts <- function(d) d$current[at_level(d, level)]

  # Assemble
  # One metric name whatever the level, so the pooled comparison keeps all
  # phenotypes; the level itself is recorded in the provenance table.
  sd_name <- "Noise SD (pA)"
  M <- function(metric, method, direction, value)
    tibble::tibble(phenotype = ph, metric = metric, method = method,
                   direction = direction, value = as.numeric(value))

  # Per-test equivalence detail (Delta, upper bound, n, verdict). Kept out of
  # `rows`; filled only from values already computed (no extra random draws).
  eqd <- list()
  EQ <- function(test, arm, delta, eq = NULL, note = NA_character_) {
    eqd[[length(eqd) + 1]] <<- tibble::tibble(
      phenotype = ph, test = test, arm = arm, delta = as.numeric(delta),
      upper = if (is.null(eq$upper)) NA_real_ else eq$upper,
      p     = if (is.null(eq$p))     NA_real_ else eq$p,
      n_real = if (is.null(eq$n_real)) NA_real_ else as.numeric(eq$n_real),
      n_arm  = if (is.null(eq$n_arm))  NA_real_ else as.numeric(eq$n_arm),
      equivalent = if (is.null(eq$equivalent)) NA else eq$equivalent,
      note = note)
  }

  message("  [", ph, " 1/6] amplitude window equivalence (real floor)...")
  amp_delta <- amplitude_floor_windows(real$current)

  # Parametric amplitude equivalence (two-Gaussian fit, fit_two_gaussian_em).
  # Mode is set once per phenotype from the real point estimate: if real does
  # not separate into two Gaussians, every fit for the phenotype uses the
  # single-Beta fit (fit_beta_single). Real bootstrap fits are shared across
  # arms and parameters.
  param_polarity <- POL_MAIN[[ph]]
  param_mode <- if (fit_two_gaussian_em(real$current, param_polarity)$ok) "gaussian" else "beta"
  param_beta_range <- if (param_mode == "beta") {
    pad <- 0.01 * diff(range(real$current))
    range(real$current) + c(-pad, pad)
  } else NULL
  if (param_mode == "beta") {
    message(ph, ": two-Gaussian EM did not separate on real; ",
            "using the single-component Beta fit (location/spread only).")
  }
  message("  [", ph, " 2/6] parametric TOST (real floor fits, mode=", param_mode, ")...")
  param_floor  <- param_floor_fits(real$current, real$channels, param_polarity,
                                    param_mode, param_beta_range)
  # Gaussian-mode sd and weight are tested by the noise-SD and Popen
  # distribution tests below. Beta mode has neither.
  PARAM_LABELS <- if (param_mode == "gaussian") {
    c(baseline = "Baseline (pA)", amplitude = "Unitary amplitude (pA)")
  } else {
    c(location = "Location (Beta fit, pA)", spread = "Spread (Beta fit, pA)")
  }
  # Gaussian mode: baseline/amplitude Delta is anchored to real's noise SD
  # (noise_sd_delta_for_param). Beta mode: location/spread use the bootstrap
  # Delta.
  param_delta <- setNames(
    lapply(names(PARAM_LABELS), function(f) {
      if (param_mode == "gaussian") {
        noise_sd_delta_for_param(f, real$current, real$channels)
      } else {
        param_delta_field(param_floor, f)
      }
    }),
    names(PARAM_LABELS))

  # Baseline/amplitude read from the channels column
  # (channels_baseline_amplitude); runs for either param_mode. Same noise-SD
  # Delta as the Gaussian-mode fit (noise_sd_delta_for_param).
  CHAN_AMP_FIELDS <- c(baseline = "Baseline (from channels, pA)",
                       amplitude = "Unitary amplitude (from channels, pA)")
  chan_amp_delta <- setNames(
    lapply(names(CHAN_AMP_FIELDS), function(f)
      noise_sd_delta_for_param(f, real$current, real$channels)),
    names(CHAN_AMP_FIELDS))
  message("  [", ph, " 3/6] channels-direct amplitude equivalence (real reps)...")
  chan_amp_real_pt   <- channels_baseline_amplitude(real$current, real$channels)
  chan_amp_real_reps <- channels_amp_real_reps(real$current, real$channels)

  # Same baseline/amplitude as a percentage of real's amplitude
  # (channels_amp_relative_delta).
  CHAN_AMP_REL_FIELDS <- c(baseline = "Baseline (relative, %)",
                           amplitude = "Unitary amplitude (relative, %)")
  chan_amp_rel_delta <- if (chan_amp_real_pt$ok) {
    setNames(
      lapply(names(CHAN_AMP_REL_FIELDS), function(f)
        channels_amp_relative_delta(f, real$current, real$channels,
                                     chan_amp_real_pt$amplitude)),
      names(CHAN_AMP_REL_FIELDS))
  } else {
    setNames(lapply(names(CHAN_AMP_REL_FIELDS), function(f) list(ok = FALSE, n = 0)),
             names(CHAN_AMP_REL_FIELDS))
  }

  # Noise-SD distribution equivalence per qualifying open/closed event
  # (noise_sd_by_event), one-sided (Wellek 2010). Runs for either param_mode.
  message("  [", ph, " 4/6] noise-SD distribution equivalence (per-event, real floor)...")
  NOISE_SD_STATES <- c("Open", "Closed")
  noise_sd_real_by_state <- setNames(
    lapply(NOISE_SD_STATES, function(s) noise_sd_by_event(real$current, real$channels, s)),
    NOISE_SD_STATES)
  noise_sd_delta_by_state <- setNames(
    lapply(noise_sd_real_by_state, dwell_delta),
    NOISE_SD_STATES)

  # Popen distribution equivalence per CFG$amp_win-sample window, one-sided
  # (Wellek 2010); see popen_delta/popen_equiv.
  message("  [", ph, " 5/6] Popen distribution equivalence (windowed, real floor)...")
  popen_delta_val <- popen_delta(real$channels)

  # Dwell-time distribution equivalence, one-sided, on the full open/closed
  # dwell-time distributions (dwell_delta/dwell_equiv).
  message("  [", ph, " 6/6] dwell-time distribution equivalence (real floor)...")
  DWELL_STATES  <- c("Open", "Closed")
  dwell_delta_by_state <- setNames(
    lapply(DWELL_STATES, function(s)
      dwell_delta(dw$dwell_ms[dw$source == "Real" & dw$state == s])),
    DWELL_STATES)

  rows <- dplyr::bind_rows(
    M("MMD vs real",          "Real floor",  "reference", mmd_rbf(W_r1, W_r2, gamma = gam)),
    M("MMD vs real",          "White noise", "reference", mmd_rbf(W$Real, W_wn, gamma = gam)),
    M("Amplitude equiv margin (Delta)", "Real (window floor)", "reference", amp_delta),
    M(sd_name,                "Real",        "reference", sd(level_pts(real))),
    M("Envelope SD (pA)",     "Real",        "reference", sd(env$Real)),
    M("Open probability",     "Real",        "reference", mean(real$channels > 0)),
    M("Mean open dwell (ms)", "Real",        "reference", mean_dwell("Real", "Open")),
    M("Mean closed dwell (ms)", "Real",      "reference", mean_dwell("Real", "Closed")),
    purrr::imap_dfr(PARAM_LABELS, function(label, f) {
      d <- param_delta[[f]]
      if (!d$ok) {
        fit_desc <- if (param_mode == "gaussian") "two-Gaussian" else "single-Beta"
        message(ph, ": '", label, "' ", fit_desc, " floor fit failed in too many ",
                "real block-bootstrap replicates (", d$n, "/",
                CFG$param_equiv_boot, " ok); Delta/TOST skipped.")
        return(M(paste0(label, " equiv margin (Delta)"), "skip", "reference", 0)[0, ])
      }
      M(paste0(label, " equiv margin (Delta)"), "Real (block-bootstrap fit)",
        "reference", d$delta)
    }),
    purrr::imap_dfr(CHAN_AMP_FIELDS, function(label, f) {
      d <- chan_amp_delta[[f]]
      if (!d$ok) {
        message(ph, ": '", label, "' has too few qualifying real events (", d$n,
                ") for a noise-SD Delta; skipped.")
        return(M(paste0(label, " equiv margin (Delta)"), "skip", "reference", 0)[0, ])
      }
      M(paste0(label, " equiv margin (Delta)"), "Real (noise SD)", "reference", d$delta)
    }),
    purrr::imap_dfr(CHAN_AMP_REL_FIELDS, function(label, f) {
      d <- chan_amp_rel_delta[[f]]
      if (!d$ok) {
        message(ph, ": '", label, "' has too few qualifying real events or a ",
                "zero/non-finite real amplitude; skipped.")
        return(M(paste0(label, " equiv margin (Delta)"), "skip", "reference", 0)[0, ])
      }
      M(paste0(label, " equiv margin (Delta)"), "Real (noise SD / amplitude)",
        "reference", 100 * d$delta)
    }),
    purrr::imap_dfr(dwell_delta_by_state, function(d, s) {
      if (!d$ok) {
        message(ph, ": too few Real ", s, " dwells (", d$n, ") for a dwell-time ",
                "equivalence floor; skipped.")
        return(M(paste0(s, " dwell equiv margin (Delta, ms)"), "skip", "reference", 0)[0, ])
      }
      M(paste0(s, " dwell equiv margin (Delta, ms)"), "Real (split-half)",
        "reference", d$delta)
    }),
    purrr::imap_dfr(noise_sd_delta_by_state, function(d, s) {
      if (!d$ok) {
        message(ph, ": too few qualifying Real ", s, " events (", d$n,
                ", need CFG$event_min_len = ", CFG$event_min_len, "+ samples) for a ",
                "noise-SD equivalence floor; skipped.")
        return(M(paste0(s, " noise SD equiv margin (Delta, pA)"), "skip", "reference", 0)[0, ])
      }
      M(paste0(s, " noise SD equiv margin (Delta, pA)"), "Real (split-half by event)",
        "reference", d$delta)
    }),
    if (!popen_delta_val$ok) {
      message(ph, ": too few real windows (", popen_delta_val$n, ") for a Popen ",
              "equivalence floor; skipped.")
      M("Popen equiv margin (Delta)", "skip", "reference", 0)[0, ]
    } else {
      M("Popen equiv margin (Delta)", "Real (window floor)", "reference", popen_delta_val$delta)
    },
    purrr::map_dfr(names(arms), function(a) {
      message("  [", ph, " arm:", a, "] MMD + parametric TOST fits...")
      pm <- mmd_permutation_p(W$Real, W[[a]], gamma = gam)
      param_fa    <- param_arm_fits(arms[[a]]$current, real$channels, param_polarity,
                                     param_mode, param_beta_range)
      fit_real_pt <- fit_amp_component(real$current, param_polarity, param_mode, param_beta_range)
      fit_arm_pt  <- fit_amp_component(arms[[a]]$current, param_polarity, param_mode, param_beta_range)
      dplyr::bind_rows(
        M("MMD vs real",            a, "lower",  mmd_rbf(W$Real, W[[a]], gamma = gam)),
        M("MMD permutation p",      a, "higher", pm$p),
        M("KS amplitude",           a, "lower",
          unname(ks_safe(real$current, arms[[a]]$current)$statistic)),
        {
          eq <- amplitude_equiv_windows(real$current, arms[[a]]$current, amp_delta)
          EQ("Amplitude", a, amp_delta,
             c(eq, list(n_real = CFG$n_windows_amp, n_arm = CFG$n_windows_amp)))
          dplyr::bind_rows(
            M("Amplitude equiv D upper", a, "lower", eq$upper),
            M("Amplitude equiv p",       a, "lower", eq$p)
          )
        },
        purrr::imap_dfr(PARAM_LABELS, function(label, f) {
          d <- param_delta[[f]]
          if (!d$ok) return(M(paste0(label, " TOST p"), a, "lower", 0)[0, ])
          eq <- param_equiv_field(param_floor, param_fa, f, d$delta)
          if (!eq$ok || !fit_real_pt$ok || !fit_arm_pt$ok) {
            fit_desc <- if (param_mode == "gaussian") "two-Gaussian" else "single-Beta"
            message(ph, "/", a, ": '", label, "' ", fit_desc, " fit unreliable ",
                     "(arm-side ok in ", eq$n_ok, "/", CFG$param_equiv_boot,
                     " replicates); TOST skipped.")
            return(M(paste0(label, " TOST p"), a, "lower", 0)[0, ])
          }
          dplyr::bind_rows(
            M(paste0(label, " equiv diff (point est.)"), a, "lower",
              abs(fit_real_pt[[f]] - fit_arm_pt[[f]])),
            M(paste0(label, " TOST p"), a, "lower", eq$p)
          )
        }),
        {
          chan_amp_arm_pt   <- channels_baseline_amplitude(arms[[a]]$current, arms[[a]]$channels)
          chan_amp_arm_reps <- channels_amp_arm_reps(arms[[a]]$current, arms[[a]]$channels,
                                                      real$channels)
          purrr::imap_dfr(CHAN_AMP_FIELDS, function(label, f) {
            d <- chan_amp_delta[[f]]
            if (!d$ok) return(M(paste0(label, " TOST p"), a, "lower", 0)[0, ])
            if (!chan_amp_real_pt$ok || !chan_amp_arm_pt$ok) {
              message(ph, "/", a, ": '", label, "' has no closed- or open-labelled ",
                       "samples on one side (real ok=", chan_amp_real_pt$ok, ", ", a,
                       " ok=", chan_amp_arm_pt$ok, "); skipped.")
              return(M(paste0(label, " TOST p"), a, "lower", 0)[0, ])
            }
            eq <- channels_amp_equiv(chan_amp_real_reps, chan_amp_arm_reps, f, d$delta)
            if (!eq$ok) {
              message(ph, "/", a, ": '", label, "' (from channels): too few usable ",
                       "block-bootstrap replicates (", eq$n_ok, "/",
                       CFG$param_equiv_boot, "); TOST skipped.")
              return(M(paste0(label, " TOST p"), a, "lower", 0)[0, ])
            }
            dplyr::bind_rows(
              M(paste0(label, " equiv diff (point est.)"), a, "lower",
                abs(chan_amp_real_pt[[f]] - chan_amp_arm_pt[[f]])),
              M(paste0(label, " TOST p"), a, "lower", eq$p)
            )
          })
        },
        {
          purrr::imap_dfr(CHAN_AMP_REL_FIELDS, function(label, f) {
            d <- chan_amp_rel_delta[[f]]
            if (!d$ok) return(M(paste0(label, " TOST p"), a, "lower", 0)[0, ])
            if (!chan_amp_real_pt$ok || !chan_amp_arm_pt$ok) {
              return(M(paste0(label, " TOST p"), a, "lower", 0)[0, ])
            }
            eq <- channels_amp_relative_equiv(chan_amp_real_reps, chan_amp_arm_reps, f,
                                               d$delta, chan_amp_real_pt$amplitude)
            if (!eq$ok) {
              message(ph, "/", a, ": '", label, "' had too few usable relative-diff ",
                       "replicates (", eq$n_ok, "/", CFG$param_equiv_boot,
                       "; fit-fail=", eq$n_discarded_fitfail, ", amplitude-guard=",
                       eq$n_discarded_guard, "); TOST skipped.")
              return(M(paste0(label, " TOST p"), a, "lower", 0)[0, ])
            }
            dplyr::bind_rows(
              M(paste0(label, " equiv diff (point est., %)"), a, "lower",
                100 * abs((chan_amp_real_pt[[f]] - chan_amp_arm_pt[[f]]) /
                          chan_amp_real_pt$amplitude)),
              M(paste0(label, " TOST p"), a, "lower", eq$p)
            )
          })
        },
        purrr::imap_dfr(dwell_delta_by_state, function(d, s) {
          if (!d$ok) {
            EQ(paste(s, "dwell"), a, NA, note = "skipped: too few real dwells for a margin")
            return(M(paste0(s, " dwell equiv p"), a, "lower", 0)[0, ])
          }
          arm_dwell <- dw$dwell_ms[dw$source == a & dw$state == s]
          eq <- dwell_equiv(dw$dwell_ms[dw$source == "Real" & dw$state == s],
                             arm_dwell, d$delta)
          EQ(paste(s, "dwell"), a, d$delta, eq,
             if (!eq$ok) paste0("skipped: ", length(arm_dwell), " arm dwells") else NA)
          if (!eq$ok) {
            message(ph, "/", a, ": too few ", s, " dwells (", length(arm_dwell),
                    ") for a dwell-time equivalence test; skipped.")
            return(M(paste0(s, " dwell equiv p"), a, "lower", 0)[0, ])
          }
          M(paste0(s, " dwell equiv p"), a, "lower", eq$p)
        }),
        purrr::imap_dfr(noise_sd_delta_by_state, function(d, s) {
          if (!d$ok) {
            EQ(paste(s, "noise SD"), a, NA,
               eq = list(n_real = length(noise_sd_real_by_state[[s]])),
               note = "skipped: too few qualifying real events for a margin")
            return(M(paste0(s, " noise SD equiv p"), a, "lower", 0)[0, ])
          }
          arm_sd <- noise_sd_by_event(arms[[a]]$current, arms[[a]]$channels, s)
          eq <- dwell_equiv(noise_sd_real_by_state[[s]], arm_sd, d$delta)
          EQ(paste(s, "noise SD"), a, d$delta, eq,
             if (!eq$ok) paste0("skipped: ", length(arm_sd), " qualifying arm events") else NA)
          if (!eq$ok) {
            message(ph, "/", a, ": too few qualifying ", s, " events (", length(arm_sd),
                    ") for a noise-SD equivalence test; skipped.")
            return(M(paste0(s, " noise SD equiv p"), a, "lower", 0)[0, ])
          }
          M(paste0(s, " noise SD equiv p"), a, "lower", eq$p)
        }),
        if (!popen_delta_val$ok) {
          M("Popen equiv p", a, "lower", 0)[0, ]
        } else {
          eq <- popen_equiv(real$channels, arms[[a]]$channels, popen_delta_val$delta)
          EQ("Popen", a, popen_delta_val$delta, eq,
             if (!eq$ok) paste0("skipped: ", eq$n, " arm windows") else NA)
          if (!eq$ok) {
            message(ph, "/", a, ": too few ", a, " windows (", eq$n,
                    ") for a Popen equivalence test; skipped.")
            M("Popen equiv p", a, "lower", 0)[0, ]
          } else {
            M("Popen equiv p", a, "lower", eq$p)
          }
        },
        M("log10 PSD RMS",          a, "lower",  psd_rms(a)),
        M("ACF RMS",                a, "lower",  acf_rms(a)),
        M("MMD envelope",           a, "lower",  mmd_rbf(We$Real, We[[a]])),
        M(sd_name,                  a, "match_real", sd(level_pts(arms[[a]]))),
        M("Envelope SD (pA)",       a, "match_real", sd(env[[a]])),
        M("Open probability",       a, "match_real", mean(arms[[a]]$channels > 0)),
        M("Mean open dwell (ms)",   a, "match_real", mean_dwell(a, "Open")),
        M("Mean closed dwell (ms)", a, "match_real", mean_dwell(a, "Closed"))
      )
    })
  )

  # Curves and provenance
  # time_ms is built from the sample index and the arm's assigned rate. The
  # DDPM export's own time column is a fixed 1e-4 s step.
  traces <- purrr::imap_dfr(arms_all, function(d, a) {
    n <- min(nrow(d), round(CFG$trace_secs * fs[[a]]))
    s <- best_window(d$channels, n)
    d[s:(s + n - 1), ] |>
      dplyr::mutate(source = a, phenotype = ph,
                    time_ms = 1000 * (dplyr::row_number() - 1) / fs[[a]],
                    win_start = s, win_popen = mean(channels > 0))
  })

  amp <- purrr::imap_dfr(arms_all, function(d, a)
    tibble::tibble(current = d$current, source = a, phenotype = ph))

  info <- tibble::tibble(
    phenotype = ph,
    seed_file = basename(row$raw),
    gan_file  = if (has_gan) basename(row$gan) else "none",
    ddpm_file = basename(row$ddpm),
    n_real = nrow(real),
    n_ddpm = nrow(arms$Diffusion),
    n_gan  = if (has_gan) nrow(arms$GAN) else NA_integer_,
    popen_real = mean(real$channels > 0),
    max_level_real = max(real$channels),
    fs_real = fs[["Real"]],
    fs_ddpm = fs[["Diffusion"]],
    fs_gan  = if (has_gan) fs[["GAN"]] else NA_real_,
    noise_level = level,
    psd_nperseg = nperseg,
    psd_seg_ms = 1000 * nperseg / fs[["Real"]],
    # Segments each arm could supply, then the count used. Indexed by name.
    psd_n_seg_real = unname(seg_avail["Real"]),
    psd_n_seg_ddpm = unname(seg_avail["Diffusion"]),
    psd_n_seg_gan  = if (has_gan) unname(seg_avail["GAN"]) else NA_integer_,
    psd_n_seg_used = psd$n_seg[psd$source == "Real"][1],
    psd_n_seg_ok   = min(c(seg_avail[["Real"]], seg_cap_sim)) >= CFG$psd_n_seg_min,
    acf_n_used     = acf_tbl$n_used[acf_tbl$source == "Real"][1],
    lag1_real = unname(lag1["Real"]),
    lag1_ddpm = unname(lag1["Diffusion"]),
    lag1_gan  = if (has_gan) unname(lag1["GAN"]) else NA_real_,
    gamma = gam
  )

  list(metrics = rows,
       equiv = dplyr::bind_rows(eqd),
       psd = dplyr::mutate(psd, phenotype = ph),
       acf = dplyr::mutate(acf_tbl, phenotype = ph),
       dwell = dplyr::mutate(dw, phenotype = ph),
       env = purrr::imap_dfr(env, function(e, a)
         tibble::tibble(idx = seq_along(e), value = e, source = a, phenotype = ph)),
       traces = traces,
       amp = amp,
       info = info)
}

# Cache key: segment length, equalisation, rates and policy label.
run_one <- function(row, fs, policy) {
  ph  <- row$phenotype
  nps <- NPERSEG_OF[[ph]]
  key <- sprintf("%s_seg%d_%s_%s_%s_%s", ph, nps, LEVEL_OF[[ph]], policy,
                 if (isTRUE(CFG$equalise_noise)) "eq" else "raw",
                 paste(sprintf("%s%g", substr(names(fs), 1, 1), fs), collapse = ""))
  cache_file <- file.path(CFG$out_dir, "cache", paste0(key, ".rds"))
  if (CFG$cache && file.exists(cache_file)) {
    message("Reusing cached ", key)
    return(readRDS(cache_file))
  }
  message("Analysing ", ph, " (", policy, " rates)")
  res <- analyse_one(row, recs[[ph]], fs, LEVEL_OF[[ph]], nps)
  if (CFG$cache) saveRDS(res, cache_file)
  res
}

results <- purrr::map(seq_len(nrow(manifest)), function(i) {
  row <- manifest[i, ]
  run_one(row, FS_MAIN[[row$phenotype]], CFG$rate_policy)
})
names(results) <- manifest$phenotype

pull_all <- function(what) purrr::map_dfr(results, what)

metrics <- pull_all("metrics")
info    <- pull_all("info")

# Stop on any non-finite metric; an NA would drop a phenotype from the paired
# comparison.
bad <- dplyr::filter(metrics, !is.finite(value))
if (nrow(bad)) {
  stop("Non-finite metrics:\n",
       paste(sprintf("  %s / %s / %s", bad$phenotype, bad$method, bad$metric),
             collapse = "\n"))
}
```

```{r records-analysed-table}
nice_kable(
  info |> dplyr::select(phenotype, seed_file, ddpm_file, gan_file,
                        n_real, n_ddpm, n_gan, popen_real, max_level_real),
  digits = 4, label = "records_analysed", caption = "Records analysed."
)
```

```{r psd-provenance-table}
nice_kable(info |> dplyr::select(phenotype, noise_level, fs_real, fs_ddpm,
                                   fs_gan, psd_nperseg, psd_seg_ms,
                                   psd_n_seg_real, psd_n_seg_ddpm,
                                   psd_n_seg_gan, psd_n_seg_used,
                                   psd_n_seg_ok, acf_n_used, gamma),
             digits = 4, label = "psd_provenance",
             caption = paste("Rate applied, spectral segment length, segments",
                             "each arm could supply, segments used by Real",
                             "(uncapped) and by GAN/Diffusion (shared minimum),",
                             "and kernel bandwidth. psd_n_seg_ok is FALSE where",
                             "either spectrum has too few segments."))
```

```{r lag1-table}
nice_kable(info |> dplyr::select(phenotype, lag1_real, lag1_ddpm, lag1_gan),
             digits = 4, label = "lag1_autocorr",
             caption = "Lag-1 autocorrelation of the single-level noise, per arm.")
readr::write_csv(info, file.path(CFG$out_dir, "provenance.csv"))
readr::write_csv(metrics, file.path(CFG$out_dir, "metrics_long.csv"))
readr::write_csv(pull_all("equiv"), file.path(CFG$out_dir, "equiv_details.csv"))
```

# Scoring

Score: the distance itself, 1 - p for the permutation p, and |arm - real| for
summary statistics such as the noise SD. Lower is better for all.

```{r score}
refs <- metrics |>
  dplyr::filter(direction == "reference", method == "Real") |>
  dplyr::select(phenotype, metric, ref = value)

scored <- metrics |>
  dplyr::filter(direction %in% c("lower", "higher", "match_real")) |>
  dplyr::left_join(refs, by = c("phenotype", "metric")) |>
  dplyr::mutate(
    score = dplyr::case_when(
      direction == "match_real" ~ abs(value - ref),
      direction == "higher"     ~ 1 - value,
      TRUE                      ~ value
    ),
    better = "lower"
  ) |>
  dplyr::filter(is.finite(score))


nice_kable(
  scored |> dplyr::arrange(metric, method, phenotype) |>
    dplyr::select(metric, method, phenotype, value, ref, score),
  digits = 5, caption = "Raw values and the comparable score derived from them."
)
readr::write_csv(scored, file.path(CFG$out_dir, "scores_long.csv"))
```

## Head to head, per phenotype

```{r head-to-head}
head_to_head <- scored |>
  dplyr::select(metric, phenotype, method, value) |>
  tidyr::pivot_wider(names_from = method, values_from = value) |>
  dplyr::left_join(refs, by = c("phenotype", "metric")) |>
  dplyr::rename(`Reference (real)` = ref) |>
  dplyr::arrange(metric, phenotype)

nice_kable(head_to_head, digits = 5,
             caption = paste("Similarity metrics for both methods per phenotype,",
                             "with the real reference value."))
readr::write_csv(head_to_head, file.path(CFG$out_dir, "head_to_head_metrics.csv"))
```

# Pooled comparison

Paired design: both methods use the same seed record, so the comparison is on
within-phenotype differences.

```{r paired}
pairable <- scored |>
  dplyr::filter(method %in% c("Diffusion", "GAN")) |>
  dplyr::group_by(metric, phenotype) |>
  dplyr::filter(dplyr::n_distinct(method) == 2) |>
  dplyr::ungroup()

n_ph <- dplyr::n_distinct(pairable$phenotype)

# Figure: the eleven similarity metrics only. Equivalence and TOST p-values are
# pooled below but not drawn.
SLOPE_METRICS <- c("ACF RMS", "Envelope SD (pA)", "KS amplitude", "MMD envelope",
                   "MMD permutation p", "MMD vs real", "Mean closed dwell (ms)",
                   "Mean open dwell (ms)", "Noise SD (pA)", "Open probability",
                   "log10 PSD RMS")
stopifnot(all(SLOPE_METRICS %in% pairable$metric))

p_slope <- ggplot(dplyr::filter(pairable, metric %in% SLOPE_METRICS), aes(factor(method, levels = c("Diffusion", "GAN")),
                                score, group = phenotype)) +
  geom_line(colour = "grey60", linewidth = 0.4) +
  geom_point(aes(colour = method), size = 2.2) +
  facet_wrap(~ metric, scales = "free_y") +
  scale_colour_manual(values = pal, guide = "none") +
  labs(x = NULL, y = "Score (lower is better)")+
  theme_channel()
p_slope
save_fig(p_slope, "fig_paired_metrics", width = 9, height = 6.5)
```

Each line joins the two scores for one phenotype.

```{r pooled-summary}
wide <- pairable |>
  dplyr::select(metric, phenotype, better, method, score) |>
  tidyr::pivot_wider(names_from = method, values_from = score) |>
  dplyr::mutate(diff = Diffusion - GAN)

paired_stats <- wide |>
  dplyr::group_by(metric, better) |>
  dplyr::summarise(
    n              = dplyr::n(),
    n_ties         = sum(diff == 0),
    median_diff    = median(diff),
    mean_diff      = mean(diff),
    sd_diff        = sd(diff),
    n_favour_diff  = sum(if (dplyr::first(better) == "lower") diff < 0 else diff > 0),
    p_paired_t     = if (dplyr::n() > 2 && sd(diff) > 0)
                       stats::t.test(diff)$p.value else NA_real_,
    ci_lo          = if (dplyr::n() > 2 && sd(diff) > 0)
                       stats::t.test(diff)$conf.int[1] else NA_real_,
    ci_hi          = if (dplyr::n() > 2 && sd(diff) > 0)
                       stats::t.test(diff)$conf.int[2] else NA_real_,
    # Sign test in the same tail as `better`; exact ties excluded.
    p_sign         = {
      fav <- if (dplyr::first(better) == "lower") sum(diff < 0) else sum(diff > 0)
      nz  <- sum(diff != 0)
      if (nz > 0) stats::binom.test(fav, nz)$p.value else NA_real_
    },
    .groups = "drop"
  ) |>
  dplyr::mutate(
    n_eff   = n - n_ties,
    # Unanimity from the counts (the verdict string may carry " (n tied)").
    unanimous = n_eff > 0 & (n_favour_diff == n_eff | n_favour_diff == 0),
    verdict = dplyr::case_when(
      n_eff == 0             ~ "no separation: all differences exactly tied",
      n_favour_diff == n_eff ~ "consistent: Diffusion",
      n_favour_diff == 0     ~ "consistent: GAN",
      TRUE                   ~ "inconsistent across phenotypes"),
    verdict = dplyr::if_else(n_ties > 0 & n_eff > 0,
                             paste0(verdict, " (", n_ties, " tied)"), verdict)) |>
  dplyr::select(metric, better, n, n_ties, median_diff, mean_diff, sd_diff,
                n_favour_diff, p_paired_t, ci_lo, ci_hi, p_sign, unanimous,
                verdict)

nice_kable(paired_stats, digits = 5,
             caption = paste("Pooled paired comparison, Diffusion minus",
                             "GAN within each phenotype: paired t-test on the",
                             "differences and sign test excluding exact ties."))
readr::write_csv(paired_stats, file.path(CFG$out_dir, "paired_summary.csv"))
```

## Sample size

```{r power}
min_p_sign <- min(1, 2 * 0.5^n_ph)
cat(sprintf("Phenotypes pooled: %d\n", n_ph))
cat(sprintf("Smallest two-sided p a sign test can return here: %.4f\n", min_p_sign))

if (min_p_sign > 0.05) {
  cat("\nAbove 0.05: no sign or Wilcoxon signed-rank test can reach p < 0.05 at this n.\n",
      sep = "")
} else {
  cat("\nA sign test can reach p < 0.05 at this n.\n")
}

cat(sprintf("\nUnanimous metrics: %d of %d.\n",
            sum(paired_stats$unanimous), nrow(paired_stats)))

if (any(paired_stats$n_ties > 0)) {
  cat("Metrics with exact ties (excluded from the unanimity count): ",
      paste(paired_stats$metric[paired_stats$n_ties > 0], collapse = ", "),
      ".\n", sep = "")
}
```

The paired t-test assumes roughly normal differences, which five phenotypes
cannot check.

# Publication figures

## Figure: example traces

```{r fig-traces, fig.height = 7}
traces <- pull_all("traces") |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")))

# Popen of the window shown, printed on each panel.
win_lab <- traces |>
  dplyr::group_by(phenotype, source) |>
  dplyr::summarise(win_popen = dplyr::first(win_popen),
                   win_start = dplyr::first(win_start),
                   x = min(time_ms), y = max(current), .groups = "drop") |>
  dplyr::mutate(label = sprintf("Popen %.2f (from %d)", win_popen, win_start))

p_traces <- ggplot(traces, aes(time_ms, current, colour = source)) +
  geom_line(linewidth = 0.2) +
  geom_text(data = win_lab, aes(x, y, label = label), hjust = 0, vjust = 1,
            size = 2.4, colour = "grey30", inherit.aes = FALSE) +
  facet_grid(phenotype ~ source, scales = "free_y") +
  scale_colour_manual(values = pal, guide = "none") +
  labs(x = "Time (ms)", y = "Current (pA)") +
  theme_channel()
p_traces
save_fig(p_traces, "fig_traces", width = 8, height = 1.4 * n_ph + 1)
```

The window is chosen by the same rule as the Python previews. Where no window
falls inside the target Popen band (Phenotype_B too quiet, Phenotype_D too
busy), the nearest is used.

## Figure: idealisations

```{r fig-ideal, fig.height = 7}
p_ideal <- ggplot(traces, aes(time_ms, channels, colour = source)) +
  geom_step(linewidth = 0.3) +
  facet_grid(phenotype ~ source) +
  scale_colour_manual(values = pal, guide = "none") +
  labs(x = "Time (ms)", y = "Open channels") +
  theme_channel()
p_ideal
save_fig(p_ideal, "fig_idealisations", width = 8, height = 1.2 * n_ph + 1)
```

## Figure: all-points amplitude histograms

```{r fig-amp, fig.height = 6}
p_amp <- pull_all("amp") |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN"))) |>
  ggplot(aes(current, colour = source, fill = source)) +
  geom_density(alpha = 0.15, linewidth = 0.5, adjust = 0.6) +
  facet_wrap(~ phenotype, scales = "free", ncol = 2) +
  scale_colour_manual(values = pal) + scale_fill_manual(values = pal) +
  labs(x = "Current (pA)", y = "Density") +
  theme_channel()
p_amp
save_fig(p_amp, "fig_amplitude", width = 8, height = 2.2 * ceiling(n_ph / 2) + 1)
```

## Figure: single-level noise

```{r fig-noise, fig.height = 6}
# Panel labels give the level and segment length used.
panel_label <- setNames(sprintf("%s (%s level, %d-sample segments)",
                                names(LEVEL_OF), LEVEL_OF,
                                NPERSEG_OF[names(LEVEL_OF)]),
                        names(LEVEL_OF))

p_psd <- pull_all("psd") |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")),
                phenotype = panel_label[phenotype]) |>
  ggplot(aes(freq, psd, colour = source)) +
  geom_line(linewidth = 0.4) +
  geom_vline(xintercept = CFG$filter_hz, linetype = "dashed",
             colour = "grey60", linewidth = 0.3) +
  facet_wrap(~ phenotype, ncol = 2) +
  scale_x_log10(labels = scales::label_number()) +
  scale_y_log10(labels = scales::label_scientific()) +
  scale_colour_manual(values = pal) +
  labs(x = "Frequency (Hz)", y = expression(PSD~(pA^2/Hz))) +
  theme_channel()

p_acf <- pull_all("acf") |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")),
                phenotype = panel_label[phenotype]) |>
  ggplot(aes(lag_ms, acf, colour = source)) +
  geom_hline(yintercept = 0, colour = "grey70", linewidth = 0.3) +
  geom_line(linewidth = 0.4) +
  facet_wrap(~ phenotype, ncol = 2) +
  scale_colour_manual(values = pal) +
  labs(x = "Lag (ms)", y = "Autocorrelation") +
  theme_channel()

p_psd
p_acf
save_fig(p_psd, "fig_psd", width = 8, height = 2.2 * ceiling(n_ph / 2) + 1)
save_fig(p_acf, "fig_acf", width = 8, height = 2.2 * ceiling(n_ph / 2) + 1)
```

## Figure: dwell times

Sigworth-Sine display: log abscissa, square-root ordinate. The diffusion arm's
`Channels` column is the conditioning idealisation it was given (known ground
truth); the GAN's labels are its own output.

```{r dwellbars,  fig.height = 7}
d_dwell <- pull_all("dwell") |>
  dplyr::filter(dwell_ms > 0) |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")))

# Log-spaced breaks: equal bar widths on the log axis.
brk <- 10^seq(log10(min(d_dwell$dwell_ms)), log10(max(d_dwell$dwell_ms)),
              length.out = 41)

p_dwell <- ggplot(d_dwell, aes(dwell_ms, colour = source)) +
  geom_histogram(aes(y = after_stat(sqrt(count)), fill = source),
                 breaks = brk, alpha = 0.15, position = "identity", colour = NA) +
  geom_histogram(aes(y = after_stat(sqrt(count))), fill = NA,
                 breaks = brk, linewidth = 0.4, position = "identity") +
  facet_grid(phenotype ~ state, 
             scales = "free_y",
             labeller = labeller(
             phenotype = function(x) gsub("_", "\n", x, fixed = TRUE)
    )) +
  scale_x_log10(labels = scales::label_number()) +
  scale_colour_manual(values = pal) +
  scale_fill_manual(values = pal, guide = "none") +
  labs(x = "Dwell time (ms)", y = expression(sqrt(count))) +
  theme_channel()+
  theme(
    strip.text.y = element_text(
      angle = 270,
      hjust = 0.5,
      vjust = 0.5
    )
  ) +
    guides(colour = guide_legend(override.aes = list(
    fill = scales::alpha(pal[c("Real", "Diffusion", "GAN")], 0.15),
    linewidth = 0.4)))
p_dwell

save_fig(p_dwell, "fig_dwell", width = 8, height = 1.5 * n_ph + 1)
```



## Bursts

```{r bursts, fig.height = 6}
# Two-component exponential mixture on the closed times by EM, then the
# Colquhoun-Sakmann equal-areas tcrit: the gap length at which as many brief
# closures are misread as between-burst as long ones are misread as within.
fit_tcrit <- function(closed_ms, n_iter = 200) {
  x <- closed_ms[is.finite(closed_ms) & closed_ms > 0]
  if (length(x) < 50) return(NA_real_)
  tau <- stats::quantile(x, c(0.25, 0.9), names = FALSE)
  if (!(tau[2] > tau[1])) return(NA_real_)
  a <- c(0.5, 0.5)
  for (i in seq_len(n_iter)) {
    d1 <- a[1] * stats::dexp(x, 1 / tau[1])
    d2 <- a[2] * stats::dexp(x, 1 / tau[2])
    w  <- d1 / (d1 + d2); w[!is.finite(w)] <- 0.5
    a   <- c(mean(w), 1 - mean(w))
    tau <- c(sum(w * x) / sum(w), sum((1 - w) * x) / sum(1 - w))
    if (any(!is.finite(tau)) || any(tau <= 0)) return(NA_real_)
  }
  o <- order(tau); tau <- tau[o]; a <- a[o]
  # One closed component: no bursts, no tcrit.
  if (tau[2] / tau[1] < 3 || min(a) < 0.02) return(NA_real_)
  f <- function(t) a[1] * exp(-t / tau[1]) - a[2] * (1 - exp(-t / tau[2]))
  lo <- tau[1] * 1e-3; hi <- tau[2]
  if (!is.finite(f(lo)) || !is.finite(f(hi)) || f(lo) * f(hi) > 0) return(NA_real_)
  stats::uniroot(f, c(lo, hi), tol = 1e-9)$root
}

burst_of <- function(dw, tcrit) {
  if (!is.finite(tcrit)) return(NULL)
  sep <- dw$state == "Closed" & dw$dwell_ms >= tcrit
  b   <- cumsum(sep)
  d <- dw[!sep, , drop = FALSE]; b <- b[!sep]
  if (!nrow(d)) return(NULL)
  # First and last bursts are truncated by the record ends; dropped.
  inner <- b > min(b) & b < max(b)
  d <- d[inner, , drop = FALSE]; b <- b[inner]
  if (!nrow(d)) return(NULL)
  tibble::tibble(burst = b, state = d$state, ms = d$dwell_ms) |>
    dplyr::group_by(burst) |>
    dplyr::summarise(length_ms = sum(ms),
                     n_open    = sum(state == "Open"),
                     open_ms   = sum(ms[state == "Open"]),
                     n_gap     = sum(state == "Closed"),
                     gap_ms    = sum(ms[state == "Closed"]),
                     .groups = "drop") |>
    dplyr::filter(n_open > 0)
}

res_burst <- purrr::map(manifest$phenotype, function(ph) {
  arms <- recs[[ph]]; fs <- FS_MAIN[[ph]]
  dwl <- lapply(names(arms), function(a) dwell_times(arms[[a]]$channels, fs[[a]]))
  names(dwl) <- names(arms)
  # One tcrit per phenotype, from the real record, applied to every arm.
  tcrit <- fit_tcrit(dwl$Real$dwell_ms[dwl$Real$state == "Closed"])
  per_burst <- purrr::map_dfr(names(arms), function(a) {
    b <- burst_of(dwl[[a]], tcrit)
    if (is.null(b) || !nrow(b)) return(NULL)
    dplyr::mutate(b, phenotype = ph, source = a)
  })
  if (!nrow(per_burst)) return(list(summary = NULL, per_burst = NULL))
  list(
    per_burst = per_burst,
    summary = per_burst |>
      dplyr::group_by(phenotype, source) |>
      dplyr::summarise(
        tcrit_ms = tcrit, n_bursts = dplyr::n(),
        `Burst length (ms)`       = mean(length_ms),
        `Openings per burst`      = mean(n_open),
        `Open time in burst (ms)` = sum(open_ms) / sum(n_open),
        `Gap in burst (ms)`       = if (sum(n_gap)) sum(gap_ms) / sum(n_gap) else NA_real_,
        `Burst Popen`             = sum(open_ms) / sum(length_ms),
        .groups = "drop") |>
      dplyr::mutate(`Bursts per s` = purrr::map_dbl(source, function(a)
        sum(per_burst$source == a) / (nrow(arms[[a]]) / fs[[a]]))))
})

burst      <- purrr::map_dfr(res_burst, "summary")
bursts_raw <- purrr::map_dfr(res_burst, "per_burst")
```

```{r burst-kinetics-table}
nice_kable(dplyr::select(burst, phenotype, source, tcrit_ms, n_bursts,
                           dplyr::everything()), digits = 4,
             label = "burst_kinetics",
             caption = paste("Burst kinetics. tcrit is fitted on the",
                             "real record of each phenotype and applied to all",
                             "three arms; a phenotype whose closed times are",
                             "monoexponential yields no tcrit and drops out."))
```

```{r burst-error-table}
burst_cmp <- burst |>
  dplyr::select(-tcrit_ms, -n_bursts) |>
  tidyr::pivot_longer(-c(phenotype, source), names_to = "metric") |>
  tidyr::pivot_wider(names_from = source, values_from = value) |>
  dplyr::filter(is.finite(Real), is.finite(Diffusion), is.finite(GAN)) |>
  dplyr::mutate(err_diffusion = abs(Diffusion - Real),
                err_gan       = abs(GAN - Real),
                favours = dplyr::case_when(
  err_diffusion < err_gan ~ "Diffusion",
  err_gan < err_diffusion ~ "GAN",
  TRUE                    ~ "Tie"
  ))

nice_kable(burst_cmp, digits = 4, label = "burst_error",
             caption = "Burst metrics, absolute error against the real record.")
```

```{r burst-unanimity-table}
nice_kable(
  burst_cmp |>
    dplyr::group_by(metric) |>
    dplyr::summarise(n = dplyr::n(),
                     n_favour_diffusion = sum(favours == "Diffusion"),
                     .groups = "drop"),
  label = "burst_unanimity",
  caption = "Unanimity of the burst metrics across phenotypes.")

readr::write_csv(burst,     file.path(CFG$out_dir, "burst_kinetics.csv"))
readr::write_csv(burst_cmp, file.path(CFG$out_dir, "burst_comparison.csv"))
```

## Burst figures

```{r burst_bars,  fig.height = 7}
d_burst <- bursts_raw |>
  dplyr::transmute(phenotype, source,
                   `Burst length (ms)` = length_ms,
                   `Openings per burst` = as.numeric(n_open)) |>
  tidyr::pivot_longer(-c(phenotype, source), names_to = "quantity") |>
  dplyr::filter(value > 0) |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")))

# Log-spaced breaks per quantity (burst length and opening count span
# different ranges).
log_breaks <- function(v, n = 41) {
  r <- range(v)
  if (r[1] == r[2]) return(NULL)
  10^seq(log10(r[1]), log10(r[2]), length.out = n)
}
brk_burst <- split(d_burst$value, d_burst$quantity) |> lapply(log_breaks)

layer_hist <- function(q) {
  d <- dplyr::filter(d_burst, quantity == q)
  list(
    geom_histogram(data = d, aes(y = after_stat(sqrt(count)), fill = source),
                   breaks = brk_burst[[q]], alpha = 0.15,
                   position = "identity", colour = NA),
    geom_histogram(data = d, aes(y = after_stat(sqrt(count))), fill = NA,
                   breaks = brk_burst[[q]], linewidth = 0.4,
                   position = "identity")
  )
}


p_burst <- ggplot(d_burst, aes(value, colour = source)) +
  lapply(names(brk_burst), layer_hist) +
  facet_grid(
    phenotype ~ quantity,
    scales = "free",
    labeller = labeller(
      phenotype = function(x) gsub("_", "\n", x, fixed = TRUE)
    )
  ) +
  scale_x_log10(labels = scales::label_number()) +
  scale_colour_manual(values = pal) +
  scale_fill_manual(values = pal, guide = "none") +
  labs(
    x = NULL,
    y = expression(sqrt(count))
  ) +
  theme_channel() +
  theme(
    strip.text.y = element_text(
      angle = 270,
      hjust = 0.5,
      vjust = 0.5
    )
  ) +
  guides(
    colour = guide_legend(
      override.aes = list(
        fill = scales::alpha(
          pal[c("Real", "Diffusion", "GAN")],
          0.15
        ),
        linewidth = 0.4
      )
    )
  )


p_burst


save_fig(p_burst, "fig_burst", width = 8, height = 1.5 * n_ph + 1)
readr::write_csv(bursts_raw, file.path(CFG$out_dir, "bursts_per_burst.csv"))
```



## Label self-consistency

Each arm's current is idealised on its own with a half-amplitude criterion and
compared with that arm's `Channels` column. For the GAN this checks its labels
against its own current; for the diffusion arm, whether the current followed
the conditioning. The real record gives the floor (expert idealisation against
half-amplitude).

```{r self-consistency}
res_selfcon <- purrr::map(manifest$phenotype, function(ph) {

  arms <- recs[[ph]]
  fs   <- FS_MAIN[[ph]]

  # First pass: try unsupervised EM on every arm.
  em_rec <- purrr::map(names(arms), function(a) {

    d <- arms[[a]]

    n_lev <- max(
      1,
      max(d$channels, na.rm = TRUE)
    )

    idealise_half_amplitude(
      d$current,
      n_levels = n_lev,
      polarity = POL_MAIN[[ph]]
    )
  })

  names(em_rec) <- names(arms)

  # If EM fails for any arm, use label-anchored CV for every
  # arm in this phenotype, keeping the comparison consistent.
  em_failed <- vapply(
    em_rec,
    function(x) all(is.na(x)),
    logical(1)
  )

  phenotype_fallback <- any(em_failed)

  if (phenotype_fallback) {
    failed_arms <- names(arms)[em_failed]

    method_note <- paste0(
      "Phenotype-wide fallback because unsupervised EM failed for: ",
      paste(failed_arms, collapse = ", ")
    )
  } else {
    method_note <- ""
  }

  # Second pass: calculate self-consistency using the method
  # selected for this entire phenotype.
  out <- purrr::map(names(arms), function(a) {

    d <- arms[[a]]

    n_lev <- max(
      1,
      max(d$channels, na.rm = TRUE)
    )

    lab <- d$channels > 0

    if (phenotype_fallback) {

      rec <- idealise_label_anchored_cv(
        current = d$current,
        labels = d$channels,
        n_levels = n_lev,
        polarity = POL_MAIN[[ph]],
        block_len = 1000L,
        trim = 0.10,
        min_state_n = 50L
      )

      ideal_method <- "Phenotype-wide label-anchored CV fallback"

    } else {

      rec <- em_rec[[a]]
      ideal_method <- "Unsupervised EM"
    }

    # This occurs only if the selected phenotype-wide
    # method still fails for this arm.
    if (all(is.na(rec))) {

      failure_reason <- attr(rec, "why")

      if (is.null(failure_reason)) {
        failure_reason <- "no reason was recorded"
      }

      return(list(
        summary = tibble::tibble(
          phenotype = ph,
          source = a,
          kappa = NA_real_,
          f1_open = NA_real_,
          popen_labels = mean(lab),
          popen_recovered = NA_real_,
          open_ms_labels = NA_real_,
          open_ms_recovered = NA_real_,
          flicker = NA,
          ideal_method = ideal_method,
          note = paste(
            method_note,
            "Selected method failed:",
            failure_reason
          )
        ),
        dwell = NULL
      ))
    }

    op <- rec > 0

    tp <- sum(lab & op)
    fp <- sum(!lab & op)
    fn <- sum(lab & !op)

    dl <- dwell_times(
      as.integer(lab),
      fs[[a]]
    )

    dr <- dwell_times(
      as.integer(op),
      fs[[a]]
    )

    mo <- function(x) {

      v <- x$dwell_ms[x$state == "Open"]

      if (length(v)) {
        mean(v)
      } else {
        NA_real_
      }
    }

    list(
      summary = tibble::tibble(
        phenotype = ph,
        source = a,
        kappa = cohen_kappa(lab, op),
        f1_open = if (2 * tp + fp + fn > 0) {
          2 * tp / (2 * tp + fp + fn)
        } else {
          NA_real_
        },
        popen_labels = mean(lab),
        popen_recovered = mean(op),
        open_ms_labels = mo(dl),
        open_ms_recovered = mo(dr),
        flicker = length(rle(op)$lengths) > 0.5 * length(op),
        ideal_method = ideal_method,
        note = method_note
      ),

      dwell = dplyr::bind_rows(
        dplyr::mutate(
          dl,
          origin = "Own labels"
        ),
        dplyr::mutate(
          dr,
          origin = "Recovered from current"
        )
      ) |>
        dplyr::mutate(
          phenotype = ph,
          source = a
        )
    )
  })

  list(
    summary = purrr::map_dfr(out, "summary"),
    dwell   = purrr::map_dfr(out, "dwell")
  )
})

self_consist <- purrr::map_dfr(res_selfcon, "summary")
sc_dwell     <- purrr::map_dfr(res_selfcon, "dwell")

if (any(self_consist$flicker %in% TRUE)) {
  warning("The recovered idealisation switches on more than half the samples ",
          "in: ", paste(sprintf("%s/%s", self_consist$phenotype[self_consist$flicker %in% TRUE],
                                self_consist$source[self_consist$flicker %in% TRUE]),
                        collapse = ", "),
          ". Probably a wrong level estimate; kappa unreliable for these rows.")
}

# Recovered Popen ~ 1 - labelled Popen with negative kappa: inverted
# idealisation.
inverted <- self_consist |>
  dplyr::filter(is.finite(kappa),
                abs(popen_recovered - (1 - popen_labels)) < 0.05,
                kappa < 0)
if (nrow(inverted)) {
  warning("The idealisation looks inverted in: ",
          paste(sprintf("%s/%s", inverted$phenotype, inverted$source),
                collapse = ", "), ". Check POL_MAIN for those phenotypes.")
}

nice_kable(self_consist, digits = 4, label = "self_consistency",
             caption = paste("Label self-consistency: each arm's",
                             "Channels column against a half-amplitude",
                             "idealisation of its own current."))
```

```{r self-consistency-deficit-table}
sc_cmp <- self_consist |>
  dplyr::select(phenotype, source, kappa) |>
  tidyr::pivot_wider(names_from = source, values_from = kappa) |>
  dplyr::filter(
    is.finite(Real),
    is.finite(Diffusion),
    is.finite(GAN)
  ) |>
  dplyr::mutate(
  deficit_diffusion = Real - Diffusion,
  deficit_gan       = Real - GAN,

  error_diffusion = abs(Diffusion - Real),
  error_gan       = abs(GAN - Real),

  pct_error_diffusion = 100 * error_diffusion / abs(Real),
  pct_error_gan       = 100 * error_gan / abs(Real),

  favours = dplyr::case_when(
    error_diffusion < error_gan ~ "Diffusion",
    error_gan < error_diffusion ~ "GAN",
    TRUE                        ~ "Tie"
  )
)

nice_kable(sc_cmp, digits = 4, label = "self_consistency_deficit",
             caption = paste("Kappa deficit of each arm against the real",
                             "record's kappa, with absolute and percentage error."))

cat(sprintf("\nPhenotypes where the diffusion arm is the more self-consistent: %d of %d.\n",
            sum(sc_cmp$favours == "Diffusion"), nrow(sc_cmp)))

readr::write_csv(self_consist, file.path(CFG$out_dir, "self_consistency.csv"))
readr::write_csv(sc_cmp, file.path(CFG$out_dir, "self_consistency_comparison.csv"))
```

The gap between the two dwell distributions below is the missed-event loss:
brief openings and closures a half-amplitude criterion cannot recover from the
current.

```{r fig-selfcon, fig.width = 10, fig.height = 7}
pal_origin <- c("Own labels" = "#1b1b1b",
                "Recovered from current" = "#c0392b")

d_sc <- sc_dwell |>
  dplyr::filter(dwell_ms > 0) |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN")),
                origin = factor(origin, levels = names(pal_origin)))

brk_sc <- 10^seq(log10(min(d_sc$dwell_ms)), log10(max(d_sc$dwell_ms)),
                 length.out = 41)

p_selfcon <- ggplot(d_sc, aes(dwell_ms, colour = origin)) +
  geom_histogram(aes(y = after_stat(sqrt(count)), fill = origin),
                 breaks = brk_sc, alpha = 0.15, position = "identity",
                 colour = NA) +
  geom_histogram(aes(y = after_stat(sqrt(count))), fill = NA,
                 breaks = brk_sc, linewidth = 0.4, position = "identity") +
  facet_grid(
    phenotype ~ source + state,
    scales = "free_y",
    labeller = labeller(
      phenotype = function(x) gsub("_", "\n", x, fixed = TRUE)
    )
  ) +
  scale_x_log10(labels = scales::label_number()) +
  scale_colour_manual(values = pal_origin) +
  scale_fill_manual(values = pal_origin, guide = "none") +
  labs(x = "Dwell time (ms)", y = expression(sqrt(count))) +
  theme_channel() +
  theme(
    strip.text.y = element_text(
      angle = 270,
      hjust = 0.5,
      vjust = 0.5
    )
  ) +
  guides(
    colour = guide_legend(
      override.aes = list(
        fill = scales::alpha(pal_origin, 0.15),
        linewidth = 0.4
      )
    )
  )

p_selfcon

save_fig(p_selfcon, "fig_self_consistency", width = 10, height = 1.5 * n_ph + 1)
readr::write_csv(sc_dwell, file.path(CFG$out_dir, "self_consistency_dwells.csv"))
```

## Dwell time constants

Exponential time constants fitted to the two dwell sets from the
self-consistency step: the arm's own labels (for the diffusion arm, the
conditioning ground truth) and the half-amplitude idealisation of its current.
The difference between them is the missed-event loss.

One component is reported where two could not be supported; the mean dwell is
then the summary for that row. Closed dwells averaging three or four samples
leave no exponential structure to resolve.

```{r dwell-tau}
tau_fits <- sc_dwell |>
  dplyr::group_by(phenotype, source, state, origin) |>
  dplyr::group_modify(~ fit_exp_mixture(.x$dwell_ms)) |>
  dplyr::ungroup() |>
  dplyr::mutate(origin = dplyr::recode(origin,
                  `Own labels` = "true",
                  `Recovered from current` = "recovered"))

nice_kable(tau_fits, digits = 4,
           caption = paste("Dwell time constants. 'true' is the arm's own",
                           "label column (for the diffusion arm, the",
                           "conditioning ground truth); 'recovered' is the",
                           "idealisation of that arm's current.",
                           "n_components is 1 where a two-component fit could",
                           "not be supported."))

readr::write_csv(tau_fits, file.path(CFG$out_dir, "dwell_tau_fits.csv"))

cat(sprintf(
  "\nTwo-component fits supportable in %d of %d phenotype/state/origin cells.\n",
  sum(tau_fits$n_components == 2), nrow(tau_fits)))
```

## Figure: slow envelope

```{r fig-env, fig.height = 6}
p_env <- pull_all("env") |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN"))) |>
  ggplot(aes(value, colour = source, fill = source)) +
  geom_density(alpha = 0.15, linewidth = 0.5) +
  facet_wrap(~ phenotype, scales = "free", ncol = 2) +
  scale_colour_manual(values = pal) + scale_fill_manual(values = pal) +
  labs(x = "Envelope current (pA)", y = "Density") +
  theme_channel()
p_env
save_fig(p_env, "fig_envelope", width = 8, height = 2.2 * ceiling(n_ph / 2) + 1)
```

The density above shows the spread of the envelope. Below, the envelope as a
time course.

```{r fig-env-time, fig.height = 7}
# idx counts blocks of slow_factor samples; seconds use each arm's rate, joined
# from a lookup table.
fs_tbl <- purrr::imap_dfr(FS_MAIN, function(fs, ph)
  tibble::tibble(phenotype = ph, source = names(fs), fs = unname(fs)))

env_time <- pull_all("env") |>
  dplyr::left_join(fs_tbl, by = c("phenotype", "source")) |>
  dplyr::mutate(time_s = idx * CFG$slow_factor / fs,
                source = factor(source, levels = c("Real", "Diffusion", "GAN")))

if (any(!is.finite(env_time$time_s)))
  stop("Envelope row with no rate in FS_MAIN.")

# Display window: first CFG$env_time_secs seconds of each arm (whole record if
# shorter), cut to the shortest arm so a phenotype's panels share one time range.
env_win <- env_time |>
  dplyr::group_by(phenotype, source) |>
  dplyr::summarise(t_end = min(max(time_s), CFG$env_time_secs), .groups = "drop") |>
  dplyr::group_by(phenotype) |>
  dplyr::summarise(t_end = min(t_end), .groups = "drop")
env_time_show <- env_time |>
  dplyr::inner_join(env_win, by = "phenotype") |>
  dplyr::filter(time_s <= t_end)

p_env_time <- ggplot(env_time_show, aes(time_s, value, colour = source)) +
  geom_line(linewidth = 0.3) +
  # One y scale per phenotype row across all three arms; axis lines on every
  # panel, y labels on the Real (left) panel only.
  facet_grid2(phenotype ~ source, scales = "free_y", axes = "all",
              remove_labels = "y",
              labeller = labeller(
                phenotype = function(x) gsub("_", "\n", x, fixed = TRUE))) +
  scale_colour_manual(values = pal, guide = "none") +
  labs(x = "Time (s)", y = "Envelope current (pA)") +
  theme_channel() +
  theme(strip.text.y = element_text(angle = 270, hjust = 0.5, vjust = 0.5))

p_env_time
save_fig(p_env_time, "fig_envelope_time", width = 8, height = 1.4 * n_ph + 1)
```

## Figure: manifold projections (optional)

```{r fig-manifold, eval = CFG$do_manifold, fig.width = 7.5, fig.height = 12}
if (!all(vapply(c("uwot", "Rtsne"), requireNamespace, logical(1), quietly = TRUE))) {
  stop("Set CFG$do_manifold = FALSE or install uwot and Rtsne.")
}

proj <- purrr::map_dfr(seq_len(nrow(manifest)), function(i) {
  row <- manifest[i, ]
  d <- list(Real = read_record(row$raw), Diffusion = read_record(row$ddpm))
  if (!is.na(row$gan)) d$GAN <- read_record(row$gan)
  Wp <- lapply(d, function(x) make_windows(x$current, CFG$proj_win, CFG$n_windows_proj))
  X <- do.call(rbind, Wp)
  lab <- rep(names(Wp), vapply(Wp, nrow, integer(1)))
  X <- scale(X); X[!is.finite(X)] <- 0
  um <- uwot::umap(X, n_neighbors = 15, min_dist = 0.1, n_components = 2)
  ts <- Rtsne::Rtsne(X, dims = 2, perplexity = 30, check_duplicates = FALSE,
                     pca = TRUE)$Y
  dplyr::bind_rows(
    tibble::tibble(x = um[, 1], y = um[, 2], source = lab, method = "UMAP"),
    tibble::tibble(x = ts[, 1], y = ts[, 2], source = lab, method = "t-SNE")
  ) |> dplyr::mutate(phenotype = row$phenotype)
})

proj <- proj |>
  dplyr::mutate(source = factor(source, levels = c("Real", "Diffusion", "GAN"))) |>
  dplyr::group_by(phenotype, method) |>
  dplyr::mutate(x = scales::rescale(x), y = scales::rescale(y)) |>
  dplyr::ungroup()

# One arm per column over a grey copy of the full cloud. Coordinates are
# rescaled per phenotype so the columns share fixed scales.
plot_proj <- function(d) {
  ggplot(d, aes(x, y)) +
    geom_point(
      data = dplyr::select(d, -source),
      colour = "grey88",
      size = 0.7
    ) +
    geom_point(
      aes(colour = source),
      size = 1.1,
      alpha = 0.85
    ) +
    facet_grid(
      phenotype ~ source,
      labeller = labeller(
        phenotype = function(x) gsub("_", "\n", x)
      )
    ) +
    scale_colour_manual(values = pal) +
    guides(
      colour = guide_legend(
        override.aes = list(size = 4, alpha = 1)
      )
    ) +
    labs(x = NULL, y = NULL) +
    theme_channel(base_size = 11) +
    theme(
      axis.text = element_blank(),
      axis.ticks = element_blank(),
      legend.text = element_text(size = 12),
      legend.key.size = unit(1.5, "lines"),
      panel.spacing = unit(0.7, "lines"),
      strip.text.y = element_text(
        angle = 270,
        hjust = 0.5,
        vjust = 0.5
      )
    )
}
p_proj_umap <- plot_proj(dplyr::filter(proj, method == "UMAP"))
p_proj_tsne <- plot_proj(dplyr::filter(proj, method == "t-SNE"))
p_proj_umap
p_proj_tsne

save_fig(p_proj_umap, "fig_manifold_umap", width = 7.5, height = 2.3 * n_ph + 1)
save_fig(p_proj_tsne, "fig_manifold_tsne", width = 7.5, height = 2.3 * n_ph + 1)
readr::write_csv(proj, file.path(CFG$out_dir, "manifold_coordinates.csv"))
```

## Cluster separation of the projections

Permutation test (ClusterSignificance, as used for DeepGANnel) of whether the
real and synthetic point clouds are separable in each embedding. A large p means
they are not separable. Run within each phenotype and projection method.

```{r cluster-test, eval = CFG$do_manifold && isTRUE(CFG$do_cluster_test)}
if (!requireNamespace("ClusterSignificance", quietly = TRUE)) {
  message("ClusterSignificance is not installed (Bioconductor):\n",
          '  BiocManager::install("ClusterSignificance")')
  cluster_tbl <- NULL
} else {
  # pcp() takes a two-column matrix and a grouping factor and returns a
  # p-value for every pairwise group comparison, so one call per phenotype
  # per method covers both Real-vs-Diffusion and Real-vs-GAN.
  cluster_one <- function(d) {
    mat <- as.matrix(d[, c("x", "y")])
    grp <- factor(as.character(d$source))
    if (dplyr::n_distinct(grp) < 2 || nrow(mat) < 10) return(NULL)
    pe <- ClusterSignificance::permute(
      mat = mat, classes = as.character(grp), iter = CFG$n_perm_cluster,
      projmethod = "pcp")
    p <- ClusterSignificance::pvalue(pe)
    tibble::tibble(comparison = names(p), p = as.numeric(p))
  }

  cluster_tbl <- proj |>
    dplyr::group_by(phenotype, method) |>
    dplyr::group_modify(~ {
      r <- tryCatch(cluster_one(.x),
                    error = function(e)
                      tibble::tibble(comparison = "failed",
                                     p = NA_real_,
                                     note = conditionMessage(e)))
      if (is.null(r)) tibble::tibble(comparison = NA_character_, p = NA_real_) else r
    }) |>
    dplyr::ungroup()

  nice_kable(cluster_tbl, digits = 4,
             caption = paste("ClusterSignificance permutation p-values for",
                             "separation between the point clouds in each",
                             "embedding."))

  readr::write_csv(cluster_tbl,
                   file.path(CFG$out_dir, "cluster_significance.csv"))

  if (any(is.na(cluster_tbl$p)))
    warning("Some cluster tests returned no p-value; see the note column.")
}
```

## Figure: generation block joins

Long diffusion records are built from overlapping blocks joined by context-carry
inpainting. Sample-to-sample differences near each join are compared with those
within blocks.

Only the diffusion arm has joins (the GAN emits its record in one pass), so this
is a single-arm check. It runs on the uncropped diffusion arm, since cropping
would move the joins.

```{r fig-seams, fig.height = 7}
step <- CFG$block_len - CFG$context_len
half <- CFG$context_len %/% 2

# Joins of one record, dropping any that fall within half a context length of
# either end, where there is not enough material on both sides to compare.
joins_of <- function(n) {
  # seq() errors when the record is shorter than one block.
  if (n <= step) return(numeric(0))
  b <- seq(step, n - 1L, by = step)
  b[b > CFG$context_len & b < (n - CFG$context_len)]
}

seam_one <- function(d, ph) {
  n <- nrow(d)
  b <- joins_of(n)
  diffs <- diff(d$current)

  if (!length(b)) {
    return(list(
      summary = tibble::tibble(phenotype = ph, n_samples = n, n_joins = 0L,
                               ks_seam = NA_real_, p_seam = NA_real_,
                               note = "single block, no joins to test"),
      trace = NULL, joins = numeric(0)))
  }

  near <- unique(unlist(lapply(b, function(i) seq(i - half, i + half))))
  near <- near[near >= 1 & near <= length(diffs)]
  far  <- setdiff(seq_along(diffs), near)

  ks <- ks_safe(diffs[near], diffs[far])

  # Drawing only: the first CFG$long_record_secs, decimated. The KS test uses
  # every sample and every join.
  fs_d   <- 1 / stats::median(diff(d$time[seq_len(min(n, 1000L))]))
  n_show <- min(n, as.integer(round(CFG$long_record_secs * fs_d)))
  dec  <- max(1L, n_show %/% 20000L)
  keep <- seq(1L, n_show, by = dec)

  list(
    summary = tibble::tibble(
      phenotype = ph, n_samples = n, n_joins = length(b),
      ks_seam = unname(ks$statistic), p_seam = unname(ks$p.value),
      note = sprintf("%d samples near a join, %d within blocks",
                     length(near), length(far))),
    trace = tibble::tibble(phenotype = ph, time_s = d$time[keep],
                           current = d$current[keep]),
    joins = d$time[b[b <= n_show]])
}

res_seam <- purrr::imap(recs_full_diffusion, function(d, ph) seam_one(d, ph))

seam_tbl   <- purrr::map_dfr(res_seam, "summary")
seam_trace <- purrr::map_dfr(res_seam, "trace")
seam_lines <- purrr::imap_dfr(res_seam, function(r, ph)
  if (length(r$joins)) tibble::tibble(phenotype = ph, time_s = r$joins) else NULL)

nice_kable(seam_tbl, digits = 4,
           caption = paste("Block joins in the diffusion arm. ks_seam compares",
                           "sample-to-sample differences within half a carried",
                           "context of a join with those within blocks."))

readr::write_csv(seam_tbl, file.path(CFG$out_dir, "seam_test.csv"))
```

```{r fig-seam-panels, fig.height = 7}
if (nrow(seam_trace)) {
  p_long <- ggplot(seam_trace, aes(time_s, current)) +
    geom_line(linewidth = 0.15, colour = pal[["Diffusion"]]) +
    { if (nrow(seam_lines))
        geom_vline(data = seam_lines, aes(xintercept = time_s),
                   linetype = "dashed", colour = "grey55", linewidth = 0.3) } +
    facet_wrap(~ phenotype, scales = "free", ncol = 1) +
    labs(x = "Time (s)", y = "Current (pA)") +
    theme_channel()
  print(p_long)
  save_fig(p_long, "fig_long_record", width = 8, height = 1.4 * n_ph + 1)
}

if (!any(seam_tbl$n_joins > 0)) {
  message("No diffusion record spans more than one block; no joins to test.")
}
```

# Summary values

Summary values for results_values.csv. Pooled values are medians across
phenotypes, with the range.

```{r placeholders}
fmt <- function(x, d = 4) formatC(x, digits = d, format = "g")

pooled <- scored |>
  dplyr::group_by(metric, method) |>
  dplyr::summarise(median = median(value), lo = min(value), hi = max(value),
                   .groups = "drop")

placeholders <- dplyr::bind_rows(
  tibble::tibble(placeholder = "n_phenotypes", value = as.character(n_ph),
                 note = "paired seed records"),
  # Total seed samples, and the subset scored after cropping each phenotype to
  # its shortest arm.
  tibble::tibble(placeholder = "n_samples_total",
                 value = format(sum(seed_inventory$n_rows), big.mark = ","),
                 note = "seed samples across all phenotypes, before cropping"),
  tibble::tibble(placeholder = "n_samples_analysed",
                 value = format(sum(info$n_real), big.mark = ","),
                 note = "seed samples actually scored, after cropping each phenotype to its shortest arm"),
  # Round before unique() so floating-point error does not split one rate.
  tibble::tibble(placeholder = "sample_khz",
                 value = paste(unique(round(sort(info$fs_real) / 1000, 3)),
                               collapse = ", "),
                 note = "acquisition rate(s), kHz"),
  pooled |>
    dplyr::transmute(
      placeholder = paste0(gsub("[^A-Za-z0-9]+", "_", tolower(metric)), "_",
                           tolower(method)),
      value = sprintf("%s (%s to %s)", fmt(median), fmt(lo), fmt(hi)),
      note = paste("median across phenotypes, range in brackets")),
  paired_stats |>
    dplyr::transmute(
      placeholder = paste0("unanimity_", gsub("[^A-Za-z0-9]+", "_", tolower(metric))),
      value = sprintf("%d/%d favour Diffusion", n_favour_diff, n),
      note = verdict),
  # Diffusion-arm dwell time constants (true and recovered), listed per
  # phenotype, where a two-component fit held.
  {
    tf <- tau_fits |>
      dplyr::filter(source == "Diffusion", n_components == 2L)
    if (nrow(tf)) {
      tf |>
        dplyr::mutate(ph = gsub("^Phenotype_", "", phenotype)) |>
        dplyr::group_by(state, origin) |>
        dplyr::summarise(
          value = paste(sprintf("%s %.3g/%.3g", ph, tau_fast_ms, tau_slow_ms),
                        collapse = ", "),
          n_ph = dplyr::n(), .groups = "drop") |>
        dplyr::transmute(
          placeholder = sprintf("tau_%s_%s", tolower(state), origin),
          value = value,
          note = sprintf("fast/slow tau in ms, per phenotype, %d of %d phenotypes supported two components",
                         n_ph, dplyr::n_distinct(tau_fits$phenotype)))
    } else {
      tibble::tibble(placeholder = "tau_fits", value = "NOT SUPPORTED",
                     note = "no phenotype supported a two-component dwell fit")
    }
  },
  # Cluster separation, if the optional test ran.
  {
    if (exists("cluster_tbl") && !is.null(cluster_tbl) &&
        any(is.finite(cluster_tbl$p))) {
      cp <- dplyr::filter(cluster_tbl, is.finite(p))
      tibble::tibble(
        placeholder = "cluster_p",
        value = sprintf("%s (%s to %s)", fmt(median(cp$p)),
                        fmt(min(cp$p)), fmt(max(cp$p))),
        note = sprintf("ClusterSignificance permutation p, median over %d tests; large p means not separable",
                       nrow(cp)))
    } else {
      tibble::tibble(placeholder = "cluster_p", value = "NOT RUN",
                     note = "ClusterSignificance absent or the test failed")
    }
  },
  # Seam test: median and range over phenotypes with at least one join.
  {
    s <- dplyr::filter(seam_tbl, is.finite(ks_seam))
    if (nrow(s))
      tibble::tibble(
        placeholder = "ks_seam",
        value = sprintf("%s (%s to %s)", fmt(median(s$ks_seam)),
                        fmt(min(s$ks_seam)), fmt(max(s$ks_seam))),
        note = sprintf("median across %d phenotypes, %d joins tested in total",
                       nrow(s), sum(s$n_joins)))
    else
      tibble::tibble(placeholder = "ks_seam", value = "NOT RUN",
                     note = "no phenotype spanned more than one block")
  }
)

nice_kable(placeholders, caption = "Summary values written to results_values.csv.")

readr::write_csv(placeholders, file.path(CFG$out_dir, "results_values.csv"))
```

# Files written by this run

Flags output files not written by this run.

```{r stray-check}
written <- list.files(CFG$out_dir, pattern = "\\.(csv|tiff)$")

# Stale: more than an hour older than the newest output.
info_files <- file.info(file.path(CFG$out_dir, written))
newest <- max(info_files$mtime)
stale  <- written[difftime(newest, info_files$mtime, units = "hours") > 1]

if (length(stale)) {
  warning("Files in ", CFG$out_dir, " not written by this run: ",
          paste(stale, collapse = ", "), ".")
} else {
  cat("All", length(written), "csv/tiff outputs in", CFG$out_dir,
      "were written by this run.\n")
}

# Files other than csv/tiff outputs and the cache.
others <- setdiff(list.files(CFG$out_dir), c(written, "cache"))
if (length(others))
  warning("Non-output files in ", CFG$out_dir, ": ",
          paste(others, collapse = ", "), ".")
```

# Session information

```{r session}
sessionInfo()
```
